---
title: LBT14
subtitle: Laboratory Test Results Shift Table -- Highest NCI CTCAE Grade Post-Baseline by Baseline NCI CTCAE Grade
---

------------------------------------------------------------------------

{{< include ../../test-utils/envir_hook.qmd >}}

::: panel-tabset
## Data Setup

Please note that for each variant, the `adlb` dataset needs to be filtered on correct flags like `WGRLOFL`, `WGRHIFL`, et al., otherwise the layout function will not return the correct counts.
There is an option to create a record for a lab test where no record is found at that visit.
If you specified `add_derived_type = "PHANTOM"` & `dtype_phantom_cond`, you don't have to use the `h_adsl_adlb_merge_using_worst_flag` function to preprocess your `adlb` dataset.
Otherwise please follow the pre-processing steps below before applying the layout functions.

```{r setup, message=FALSE}
#| code-fold: show

library(scda)
library(dplyr)
library(tern)

adsl <- synthetic_cdisc_dataset("latest", "adsl")
adlb <- synthetic_cdisc_dataset("latest", "adlb")

adsl <- df_explicit_na(adsl)
adlb <- df_explicit_na(adlb)

# Please note that in real clinical data, population flag like SAFFL, and parameter category like PARCAT2 needs to be
# selected properly.
adsl_f <- adsl %>% filter(SAFFL == "Y")
adlb <- adlb %>% filter(SAFFL == "Y")
```

## Standard Table (High)

Note that the worst laboratory flag (below `WGRHIFL`) must be selected appropriately in the pre-processing step.
New grouping variables `ATOXGR_GP` (post-baseline) and `BTOXGR_GP` (baseline) are created to display the correct output.

```{r variant1, test = list(result_v1 = "result")}
adlb_f <- adlb %>% filter(WGRHIFL == "Y")

# Please note the step below can be skipped if you are using DTYPE PHANTOM
adlb_out <- h_adsl_adlb_merge_using_worst_flag(adsl_f, adlb_f, worst_flag = c("WGRHIFL" = "Y"))

# Create new grouping variables ATOXGR_GP, BTOXGR_GP
adlb_out <- adlb_out %>%
  mutate(
    ATOXGR_GP = factor(case_when(
      ATOXGR %in% c(0, -1, -2, -3, -4) ~ "Not High",
      ATOXGR == 1 ~ "1",
      ATOXGR == 2 ~ "2",
      ATOXGR == 3 ~ "3",
      ATOXGR == 4 ~ "4",
      ATOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not High", "1", "2", "3", "4", "Missing"))
  ) %>%
  mutate(
    BTOXGR_GP = factor(case_when(
      BTOXGR %in% c(0, -1, -2, -3, -4) ~ "Not High",
      BTOXGR == 1 ~ "1",
      BTOXGR == 2 ~ "2",
      BTOXGR == 3 ~ "3",
      BTOXGR == 4 ~ "4",
      BTOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not High", "1", "2", "3", "4", "Missing"))
  )

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  split_rows_by(
    "PARAM",
    split_fun = drop_split_levels,
    label_pos = "topleft",
    split_label = "Parameter"
  ) %>%
  split_rows_by(
    "BTOXGR_GP",
    label_pos = "topleft",
    split_label = "    Baseline NCI-CTCAE Grade",
    indent_mod = 2L
  ) %>%
  summarize_num_patients(var = "USUBJID", .stats = c("unique_count"), unique_count_suffix = FALSE) %>%
  count_occurrences_by_grade("ATOXGR_GP", denom = "n", drop = FALSE, .indent_mods = 3L) %>%
  append_topleft("              Post-baseline NCI-CTCAE Grade") %>%
  build_table(df = adlb_out, alt_counts_df = adsl_f) %>%
  prune_table()

result
```

## Standard Table (Low)

Note that the worst laboratory flag (below `WGRLOFL`) must be selected appropriately in the pre-processing step.
New grouping variables `ATOXGR_GP` (post-baseline) and `BTOXGR_GP` (baseline) are created to display the correct output.

```{r variant2, test = list(result_v2 = "result")}
adlb_f <- adlb %>% filter(WGRLOFL == "Y")

# Please note the step below can be skipped if you are using DTYPE PHANTOM
adlb_out <- h_adsl_adlb_merge_using_worst_flag(adsl_f, adlb_f, worst_flag = c("WGRLOFL" = "Y"))

# Create new grouping variables ATOXGR_GP, BTOXGR_GP
adlb_out <- adlb_out %>%
  mutate(
    ATOXGR_GP = factor(case_when(
      ATOXGR %in% c(0, 1, 2, 3, 4) ~ "Not Low",
      ATOXGR == -1 ~ "1",
      ATOXGR == -2 ~ "2",
      ATOXGR == -3 ~ "3",
      ATOXGR == -4 ~ "4",
      ATOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not Low", "1", "2", "3", "4", "Missing"))
  ) %>%
  mutate(
    BTOXGR_GP = factor(case_when(
      BTOXGR %in% c(0, 1, 2, 3, 4) ~ "Not Low",
      BTOXGR == -1 ~ "1",
      BTOXGR == -2 ~ "2",
      BTOXGR == -3 ~ "3",
      BTOXGR == -4 ~ "4",
      BTOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not Low", "1", "2", "3", "4", "Missing"))
  )

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  split_rows_by(
    "PARAM",
    split_fun = drop_split_levels,
    label_pos = "topleft",
    split_label = "Parameter"
  ) %>%
  split_rows_by(
    "BTOXGR_GP",
    label_pos = "topleft",
    split_label = "    Baseline NCI-CTCAE Grade",
    indent_mod = 2L
  ) %>%
  summarize_num_patients(var = "USUBJID", .stats = c("unique_count"), unique_count_suffix = FALSE) %>%
  count_occurrences_by_grade("ATOXGR_GP", denom = "n", drop = FALSE, .indent_mods = 3L) %>%
  append_topleft("              Post-baseline NCI-CTCAE Grade") %>%
  build_table(df = adlb_out, alt_counts_df = adsl_f) %>%
  prune_table()

result
```

## Table Without Patients with <br/> Missing Baseline (High)

Note that missing baseline values are filtered out in the pre-processing step.

```{r variant3, test = list(result_v3 = "result")}
adlb_f <- adlb %>% filter(WGRHIFL == "Y")

# Please note the step below can be skipped if you are using DTYPE PHANTOM
adlb_out <- h_adsl_adlb_merge_using_worst_flag(adsl_f, adlb_f, worst_flag = c("WGRHIFL" = "Y"))

# Create new grouping variables ATOXGR_GP, BTOXGR_GP
adlb_out <- adlb_out %>%
  filter(BTOXGR != "<Missing>") %>%
  mutate(
    ATOXGR_GP = factor(case_when(
      ATOXGR %in% c(0, -1, -2, -3, -4) ~ "Not High",
      ATOXGR == 1 ~ "1",
      ATOXGR == 2 ~ "2",
      ATOXGR == 3 ~ "3",
      ATOXGR == 4 ~ "4",
      ATOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not High", "1", "2", "3", "4", "Missing"))
  ) %>%
  mutate(
    BTOXGR_GP = factor(case_when(
      BTOXGR %in% c(0, -1, -2, -3, -4) ~ "Not High",
      BTOXGR == 1 ~ "1",
      BTOXGR == 2 ~ "2",
      BTOXGR == 3 ~ "3",
      BTOXGR == 4 ~ "4"
    ), levels = c("Not High", "1", "2", "3", "4"))
  )

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  split_rows_by(
    "PARAM",
    split_fun = drop_split_levels,
    label_pos = "topleft",
    split_label = "Parameter"
  ) %>%
  split_rows_by(
    "BTOXGR_GP",
    label_pos = "topleft",
    split_label = "    Baseline NCI-CTCAE Grade",
    indent_mod = 2L
  ) %>%
  summarize_num_patients(var = "USUBJID", .stats = c("unique_count"), unique_count_suffix = FALSE) %>%
  count_occurrences_by_grade("ATOXGR_GP", denom = "n", drop = FALSE, .indent_mods = 3L) %>%
  append_topleft("              Post-baseline NCI-CTCAE Grade") %>%
  build_table(df = adlb_out, alt_counts_df = adsl_f) %>%
  prune_table()

result
```

## Table with Missing Baseline <br/> Considered as Grade 0 (Low)

Note that when BTOXGR is missing, the grouping variable `BTOXGR_GP` now is `"Not Low"` instead of `"Missing"` compared to *Standard Table (Low)*.

```{r variant4, test = list(result_v4 = "result")}
adlb_f <- adlb %>% filter(WGRLOFL == "Y")

# Please note the step below can be skipped if you are using DTYPE PHANTOM
adlb_out <- h_adsl_adlb_merge_using_worst_flag(adsl_f, adlb_f, worst_flag = c("WGRLOFL" = "Y"))

# Create new grouping variables ATOXGR_GP, BTOXGR_GP
adlb_out <- adlb_out %>%
  mutate(
    ATOXGR_GP = factor(case_when(
      ATOXGR %in% c(0, 1, 2, 3, 4) ~ "Not Low",
      ATOXGR == -1 ~ "1",
      ATOXGR == -2 ~ "2",
      ATOXGR == -3 ~ "3",
      ATOXGR == -4 ~ "4",
      ATOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not Low", "1", "2", "3", "4"))
  ) %>%
  mutate(
    BTOXGR_GP = factor(case_when(
      BTOXGR %in% c(0, 1, 2, 3, 4, "<Missing>") ~ "Not Low",
      BTOXGR == -1 ~ "1",
      BTOXGR == -2 ~ "2",
      BTOXGR == -3 ~ "3",
      BTOXGR == -4 ~ "4"
    ), levels = c("Not Low", "1", "2", "3", "4", "Missing"))
  )

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  split_rows_by(
    "PARAM",
    split_fun = drop_split_levels,
    label_pos = "topleft",
    split_label = "Parameter"
  ) %>%
  split_rows_by(
    "BTOXGR_GP",
    label_pos = "topleft",
    split_label = "    Baseline NCI-CTCAE Grade",
    indent_mod = 2L
  ) %>%
  summarize_num_patients(var = "USUBJID", .stats = c("unique_count"), unique_count_suffix = FALSE) %>%
  count_occurrences_by_grade("ATOXGR_GP", denom = "n", drop = FALSE, .indent_mods = 3L) %>%
  append_topleft("              Post-baseline NCI-CTCAE Grade") %>%
  build_table(df = adlb_out, alt_counts_df = adsl_f) %>%
  prune_table()

result
```

## Table with Fill-In of Grades

Pre-processing is the same as *Standard Table (High)*, but in order to keep all levels, `prune_table()` is not applied.

```{r variant5, test = list(result_v5 = "result")}
adlb_f <- adlb %>% filter(WGRHIFL == "Y")

# Please note the step below can be skipped if you are using DTYPE PHANTOM
adlb_out <- h_adsl_adlb_merge_using_worst_flag(adsl_f, adlb_f, worst_flag = c("WGRHIFL" = "Y"))

# Create new grouping variables ATOXGR_GP, BTOXGR_GP
adlb_out <- adlb_out %>%
  mutate(
    ATOXGR_GP = factor(case_when(
      ATOXGR %in% c(0, -1, -2, -3, -4) ~ "Not High",
      ATOXGR == 1 ~ "1",
      ATOXGR == 2 ~ "2",
      ATOXGR == 3 ~ "3",
      ATOXGR == 4 ~ "4",
      ATOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not High", "1", "2", "3", "4", "Missing"))
  ) %>%
  mutate(
    BTOXGR_GP = factor(case_when(
      BTOXGR %in% c(0, -1, -2, -3, -4) ~ "Not High",
      BTOXGR == 1 ~ "1",
      BTOXGR == 2 ~ "2",
      BTOXGR == 3 ~ "3",
      BTOXGR == 4 ~ "4",
      BTOXGR == "<Missing>" ~ "Missing"
    ), levels = c("Not High", "1", "2", "3", "4", "Missing"))
  )

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  split_rows_by(
    "PARAM",
    split_fun = drop_split_levels,
    label_pos = "topleft",
    split_label = "Parameter"
  ) %>%
  split_rows_by(
    "BTOXGR_GP",
    label_pos = "topleft",
    split_label = "    Baseline NCI-CTCAE Grade",
    indent_mod = 2L
  ) %>%
  summarize_num_patients(var = "USUBJID", .stats = c("unique_count"), unique_count_suffix = FALSE) %>%
  count_occurrences_by_grade("ATOXGR_GP", denom = "n", drop = FALSE, .indent_mods = 3L) %>%
  append_topleft("              Post-baseline NCI-CTCAE Grade") %>%
  build_table(df = adlb_out, alt_counts_df = adsl_f)

result
```

{{< include ../../test-utils/save_results.qmd >}}

## `teal` App

```{r teal, message=FALSE, opts.label='skip_if_testing'}
#| screenshot.opts = list(delay = 16)

library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(scda)
  library(dplyr)

  ADSL <- synthetic_cdisc_dataset("latest", "adsl")
  ADLB <- synthetic_cdisc_dataset("latest", "adlb")
})
datanames <- c("ADSL", "ADLB")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADSL <- data[["ADSL"]]
ADLB <- data[["ADLB"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_t_shift_by_grade(
      label = "Grade Laboratory Abnormality Table",
      dataname = "ADLB",
      arm_var = choices_selected(
        choices = variable_choices(ADSL, subset = c("ARM", "ARMCD")),
        selected = "ARM"
      ),
      paramcd = choices_selected(
        choices = value_choices(ADLB, "PARAMCD", "PARAM"),
        selected = "ALT"
      ),
      worst_flag_var = choices_selected(
        choices = variable_choices(ADLB, subset = c("WGRLOVFL", "WGRLOFL", "WGRHIVFL", "WGRHIFL")),
        selected = c("WGRHIFL")
      ),
      worst_flag_indicator = choices_selected(
        value_choices(ADLB, "WGRLOVFL"),
        selected = "Y", fixed = TRUE
      ),
      anl_toxgrade_var = choices_selected(
        choices = variable_choices(ADLB, subset = c("ATOXGR")),
        selected = c("ATOXGR"),
        fixed = TRUE
      ),
      base_toxgrade_var = choices_selected(
        choices = variable_choices(ADLB, subset = c("BTOXGR")),
        selected = c("BTOXGR"),
        fixed = TRUE
      ),
      add_total = FALSE
    )
  ),
  filter = teal_slices(teal_slice("ADSL", "SAFFL", selected = "Y"))
)

shinyApp(app$ui, app$server)
```

{{< include ../../repro.qmd >}}
:::
