---
title: RMPT06
subtitle: Seriousness, Outcomes, Severity, Frequency with 95% CI for Risk Management Plan
---

------------------------------------------------------------------------

::: panel-tabset
## Data Setup

The `adae` and `adsl` datasets need to undergo data processing before table production.
Some new variables are added to these datasets to create the desired tables.

```{r setup, message=FALSE}
#| code-fold: show

library(tern)
library(dplyr)
library(scda)

adsl <- synthetic_cdisc_dataset("latest", "adsl")
adae <- synthetic_cdisc_dataset("latest", "adae")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adae <- df_explicit_na(
  adae,
  omit_columns = c("SMQ01NAM", "SMQ01SC", "SMQ02NAM", "SMQ02SC", "CQ01NAM", "STUDYID", "USUBJID")
)

df_max <- aggregate(as.numeric(AETOXGR) ~ USUBJID, data = adae, FUN = max, drop = FALSE)
colnames(df_max) <- c("USUBJID", "WTOXGR")

adae <- adae %>%
  left_join(df_max, by = c("USUBJID")) %>%
  mutate(
    WTOXGR = factor(WTOXGR, levels = c("1", "2", "3", "4", "5")),
    fl_ser = AESER == "Y"
  ) %>%
  mutate(
    WTOXGR = forcats::fct_recode(
      WTOXGR,
      "Grade 1" = "1",
      "Grade 2" = "2",
      "Grade 3" = "3",
      "Grade 4" = "4",
      "Grade 5" = "5"
    ),
    AEOUT = forcats::fct_recode(
      AEOUT,
      "Fatal outcome" = "FATAL",
      "Unresolved" = "NOT RECOVERED/NOT RESOLVED",
      "Recovered/Resolved" = "RECOVERED/RESOLVED",
      "Resolved with sequelae" = "RECOVERED/RESOLVED WITH SEQUELAE",
      "Recovering/Resolving" = "RECOVERING/RESOLVING",
      "Unknown outcome" = "UNKNOWN"
    )
  ) %>%
  var_relabel(
    WTOXGR = "Worst overall grade",
    fl_ser = "Number of patients with at least one serious AE"
  )

# Add AE flag to adsl.
adsl1 <- adsl %>%
  mutate(AEFL = ifelse(USUBJID %in% adae$USUBJID, TRUE, FALSE)) %>%
  var_relabel(AEFL = "At least one AE")
```

## Standard Table

Please note that by default the percentage analysis uses Wald's confidence interval with continuity correction and the default confidence interval for percentage is 95%.
To use a different method for confidence interval calculation, the method name needs to be provided to `method` argument.

```{r}
lyt_adsl <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  estimate_proportion(
    vars = "AEFL",
    method = "clopper-pearson",
    .labels = c(
      n_prop = "Number of patients with at least one adverse event",
      prop_ci = "95% CI for % of patients with at least one AE (Clopper-Pearson)"
    ),
    table_names = "est_prop"
  )

result_adsl <- build_table(lyt_adsl, df = adsl1, alt_counts_df = adsl1)

# Layout for variables from adae dataset.
lyt_adae <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_values(
    "STUDYID",
    values = as.character(unique(adae$STUDYID)),
    .stats = "count",
    .labels = c(count = "Total number of AEs"),
    table_names = "total_aes"
  ) %>%
  count_occurrences(
    "WTOXGR",
    var_labels = "Number of patients with at least one AE by worst grade",
    show_labels = "visible"
  ) %>%
  count_patients_with_flags(
    "USUBJID",
    flag_variables = var_labels(adae[, "fl_ser"])
  ) %>%
  count_occurrences(
    "AEOUT",
    denom = "n",
    var_labels = "Number of patients with at least one AE by outcome",
    show_labels = "visible"
  )

result_adae <- build_table(lyt_adae, df = adae, alt_counts_df = adsl)

# Combine tables.
col_info(result_adsl) <- col_info(result_adae)
result <- rbind(
  result_adsl,
  result_adae[1:nrow(result_adae), ]
)

result
```

## Table Selecting <br/> Sections to Display

```{r}
lyt_adsl <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  estimate_proportion(
    vars = "AEFL",
    method = "clopper-pearson",
    .labels = c(
      n_prop = "Number of patients with at least one adverse event",
      prop_ci = "95% CI for % of patients with at least one AE (Clopper-Pearson)"
    ),
    table_names = "est_prop"
  )

result_adsl <- build_table(lyt_adsl, df = adsl1, alt_counts_df = adsl1)

# Layout for variables from adae dataset.
lyt_adae <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_values(
    "STUDYID",
    values = as.character(unique(adae$STUDYID)),
    .stats = "count",
    .labels = c(count = "Total number of AEs"),
    table_names = "total_aes"
  ) %>%
  count_occurrences(
    "WTOXGR",
    var_labels = "Number of patients with at least one AE by worst grade",
    show_labels = "visible"
  ) %>%
  count_patients_with_flags(
    "USUBJID",
    flag_variables = var_labels(adae[, "fl_ser"]),
    denom = "N_col"
  )

result_adae <- build_table(lyt_adae, df = adae, alt_counts_df = adsl)

# Combine tables.
col_info(result_adsl) <- col_info(result_adae)
result <- rbind(
  result_adsl,
  result_adae[1:nrow(result_adae), ]
)

result
```

## Table with Differences

Please note that by default the percentage difference analysis also uses Wald's confidence interval with continuity correction and the default confidence interval is 95%.

```{r}
lyt_adsl <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "AEFL",
    method = "clopper-pearson",
    .labels = c(
      n_prop = "Number of patients with at least one adverse event",
      prop_ci = "95% CI for % of patients with at least one AE"
    ),
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "AEFL",
    .labels = c(
      diff = "Difference in % of patients with at least one AE",
      diff_ci = "95% CI of difference (Wald, with correction)"
    ),
    table_names = "est_prop_diff"
  )

result_adsl <- build_table(lyt_adsl, df = adsl1, alt_counts_df = adsl1)

# Layout for variables from adae dataset.
lyt_adae <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_values(
    "STUDYID",
    values = as.character(unique(adae$STUDYID)),
    .stats = "count",
    .labels = c(count = "Total number of AEs"),
    table_names = "total_aes"
  ) %>%
  count_occurrences(
    "WTOXGR",
    var_labels = "Number of patients with at least one AE by worst grade",
    show_labels = "visible"
  ) %>%
  count_patients_with_flags(
    "USUBJID",
    flag_variables = var_labels(adae[, "fl_ser"]),
    denom = "N_col"
  ) %>%
  count_occurrences(
    "AEOUT",
    denom = "n",
    var_labels = "Number of patients with at least one AE by outcome",
    show_labels = "visible"
  )

result_adae <- build_table(lyt_adae, df = adae, alt_counts_df = adsl)

# Combine tables.
col_info(result_adsl) <- col_info(result_adae)
result <- rbind(
  result_adsl,
  result_adae[1:nrow(result_adae), ]
)

result
```

## Table Modifying Alpha Level and <br/> Type of Confidence Interval

The confidence level can be changed through the `conf_level` argument.

```{r}
lyt_adsl <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "AEFL",
    conf_level = 0.90,
    method = "clopper-pearson",
    .labels = c(
      n_prop = "Number of patients with at least one adverse event",
      prop_ci = "90% CI for % of patients with at least one AE (Clopper-Pearson)"
    ),
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "AEFL",
    conf_level = 0.90,
    method = "newcombe",
    .labels = c(
      diff = "Difference in % of patients with at least one AE",
      diff_ci = "90% CI of difference (Newcombe)"
    ),
    table_names = "est_prop_diff"
  )

result_adsl <- build_table(lyt_adsl, df = adsl1, alt_counts_df = adsl1)

# Layout for variables from adae dataset.
lyt_adae <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_values(
    "STUDYID",
    values = as.character(unique(adae$STUDYID)),
    .stats = "count",
    .labels = c(count = "Total number of AEs"),
    table_names = "total_aes"
  ) %>%
  count_occurrences(
    "WTOXGR",
    var_labels = "Number of patients with at least one AE by worst grade",
    show_labels = "visible"
  ) %>%
  count_patients_with_flags(
    "USUBJID",
    flag_variables = var_labels(adae[, "fl_ser"]),
    denom = "N_col"
  ) %>%
  count_occurrences(
    "AEOUT",
    denom = "n",
    var_labels = "Number of patients with at least one AE by outcome",
    show_labels = "visible"
  )

result_adae <- build_table(lyt_adae, df = adae, alt_counts_df = adsl)

# Combine tables.
col_info(result_adsl) <- col_info(result_adae)
result <- rbind(
  result_adsl,
  result_adae[1:nrow(result_adae), ]
)

result
```

## Table by SMQ

```{r}
#| code-fold: show

# In progress
```

## Teal

```{r teal, message=FALSE}
#| code-fold: show

# In progress
```

{{< include ../../si.qmd >}}
:::
