---
title: LBT08
subtitle: Laboratory Test Results with Highest NCI CTCAE Grade at Any Time
---

------------------------------------------------------------------------

::: panel-tabset
## Data Setup

```{r setup, message=FALSE}
#| code-fold: show

library(scda)
library(dplyr)
library(tern)

adsl <- synthetic_cdisc_dataset("latest", "adsl")
adlb <- synthetic_cdisc_dataset("latest", "adlb")

adlb <- adlb %>%
  mutate(
    GRADDR = case_when(
      PARAMCD == "ALT" ~ "B",
      PARAMCD == "CRP" ~ "H",
      PARAMCD == "IGA" ~ "L"
    )
  ) %>%
  filter(SAFFL == "Y" & ONTRTFL == "Y" & GRADDR != "")

adsl <- df_explicit_na(adsl)
adlb <- df_explicit_na(adlb)

df <- h_adlb_worsen(
  adlb,
  worst_flag_low = c("WGRLOFL" = "Y"),
  worst_flag_high = c("WGRHIFL" = "Y"),
  direction_var = "GRADDR"
)

attributes(df$GRADDR) <- list("label" = "Direction of Abnormality")
```

## Standard Table

Note: The direction of the shift table for each lab test is based on metadata and NCI CTCAE specifications.
In addition, the worst laboratory flags must be selected appropriately to match the direction of abnormality.
For example, if any lab requires a shift table for both directions, then both `worst_flag_low` and `worst_flag_high` must be specified in `h_adlb_worsen`.
If all labs requires a shift table for only one direction, then the matching worst lab flag variable must be selected in `h_adlb_worsen`.

```{r variant1}
result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARMCD") %>%
  split_rows_by("PARAMCD", label_pos = "topleft", split_label = obj_label(df$PARAMCD)) %>%
  split_rows_by("GRADDR", label_pos = "topleft", split_label = obj_label(df$GRADDR)) %>%
  count_abnormal_lab_worsen_by_baseline(
    var = "ATOXGR",
    variables = list(
      id = "USUBJID",
      baseline_var = "BTOXGR",
      direction_var = "GRADDR"
    )
  ) %>%
  append_topleft("    Highest NCI CTCAE Grade") %>%
  build_table(df = df, alt_counts_df = adsl)

result
```

## Teal

```{r teal, message=FALSE}
#| code-fold: show

# In progress
```

{{< include ../../si.qmd >}}

:::
