---
title: BWG01
subtitle: Box Plot
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

::: panel-tabset
## Data Setup

```{r setup, message = FALSE}
#| code-fold: show

# generic code for all plots
library(dplyr)
library(ggplot2)
library(nestcolor)

adlb <- random.cdisc.data::cadlb
adlb <- adlb %>% filter(PARAMCD == "ALT" & AVISIT == "WEEK 2 DAY 15")

# Definition of boxplot boundaries and whiskers
five_num <- function(x, probs = c(0, 0.25, 0.5, 0.75, 1)) {
  r <- quantile(x, probs)
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}

# get outliers based on quantile
# for outliers based on IQR see coef in geom_boxplot
get_outliers <- function(x, probs = c(0.05, 0.95)) {
  r <- subset(x, x < quantile(x, probs[1]) | quantile(x, probs[2]) < x)
  if (!is.null(x)) {
    x_names <- subset(names(x), x < quantile(x, probs[1]) | quantile(x, probs[2]) < x)
    names(r) <- x_names
  }
  r
}

# create theme used for all plots
theme_bp <- theme(
  plot.title = element_text(hjust = 0),
  plot.subtitle = element_text(hjust = 0),
  plot.caption = element_text(hjust = 0),
  panel.background = element_rect(fill = "white", color = "grey50")
)

# assign fill color and outline color
fc <- "#eaeef5"
oc <- getOption("ggplot2.discrete.fill")[1]

# get plot metadata data to derive coordinates for adding annotations
bp_annos <- function(bp, color, annos = 1) {
  bp_mdat <- ggplot_build(bp)$data[[1]]
  if (annos == 1) {
    bp <- bp +
      geom_segment(data = bp_mdat, aes(
        x = xmin + (xmax - xmin) / 4, xend = xmax - (xmax - xmin) / 4,
        y = ymax, yend = ymax
      ), linewidth = .5, color = color) +
      geom_segment(data = bp_mdat, aes(
        x = xmin + (xmax - xmin) / 4, xend = xmax - (xmax - xmin) / 4,
        y = ymin, yend = ymin
      ), linewidth = .5, color = color)
  } else {
    bp <- bp +
      geom_segment(data = bp_mdat, aes(
        x = xmin + (xmax - xmin) / 4, xend = xmax - (xmax - xmin) / 4,
        y = ymax, yend = ymax
      ), linewidth = .5, color = color) +
      geom_segment(data = bp_mdat, aes(
        x = xmin + (xmax - xmin) / 4, xend = xmax - (xmax - xmin) / 4,
        y = ymin, yend = ymin
      ), linewidth = .5, color = color) +
      geom_segment(data = bp_mdat, aes(
        x = xmin, xend = xmax,
        y = middle, yend = middle
      ), colour = color, linewidth = .5)
  }
  return(bp)
}
```

## Standard Plot

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot1, test = list(plot_v1 = "plot")}
bp_1 <- ggplot(adlb, aes(x = ARMCD, y = AVAL)) +
  stat_summary(geom = "boxplot", fun.data = five_num, fill = fc, color = oc) +
  stat_summary(geom = "point", fun = mean, size = 3, shape = 8) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    subtitle = paste("Visit:", adlb$AVISIT[1]),
    caption = "The whiskers extend to the minimum and maximum values.",
    x = "Treatment Group",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

plot <- bp_annos(bp_1, oc)
plot
```

`r webr_code_labels <- c("setup", "plot1")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot Changing <br/> Whiskers

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot2, test = list(plot_v2 = "plot")}
bp_3 <- ggplot(adlb, aes(x = ARMCD, y = AVAL)) +
  stat_summary(
    geom = "boxplot", fun.data = five_num,
    fun.args = list(probs = c(0.05, 0.25, 0.5, 0.75, 0.95)), fill = fc, color = oc
  ) +
  stat_summary(geom = "point", fun = mean, size = 3, shape = 8) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    subtitle = paste("Visit:", adlb$AVISIT[1]),
    caption = "The whiskers extend to the 5th and 95th percentile. Values outside the whiskers have not been plotted.",
    x = "Treatment Group",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

plot <- bp_annos(bp_3, oc)
plot
```

`r webr_code_labels <- c("setup", "plot2")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot Adding <br/> Outliers

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot3, test = list(plot_v3 = "plot")}
bp_4 <- ggplot(adlb, aes(x = ARMCD, y = AVAL)) +
  stat_summary(
    geom = "boxplot", fun.data = five_num,
    fun.args = list(probs = c(0.05, 0.25, 0.5, 0.75, 0.95)), fill = fc, color = oc
  ) +
  stat_summary(geom = "point", fun = get_outliers, shape = 1) +
  stat_summary(geom = "point", fun = mean, size = 3, shape = 8) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    subtitle = paste("Visit:", adlb$AVISIT[1]),
    caption = "The whiskers extend to the 5th and 95th percentile.",
    x = "Treatment Group",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

plot <- bp_annos(bp_4, oc)
plot
```

`r webr_code_labels <- c("setup", "plot3")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot Specifying Marker <br/> for Outliers and <br/> Adding Patient ID

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot4, test = list(plot_v4 = "plot")}
adlb_o <- adlb %>%
  group_by(ARMCD) %>%
  mutate(OUTLIER = AVAL < quantile(AVAL, 0.01) | AVAL > quantile(AVAL, 0.99)) %>%
  filter(OUTLIER == TRUE) %>%
  select(ARMCD, AVAL, USUBJID)
# Next step may be study-specific: shorten USUBJID to make annotation labels
# next 2 lines of code split USUBJID by "-" and take the last syllable as ID
n_split <- max(vapply(strsplit(adlb_o$USUBJID, "-"), length, numeric(1)))
adlb_o$ID <- vapply(strsplit(adlb_o$USUBJID, "-"), `[[`, n_split, FUN.VALUE = "a")

bp_5 <- ggplot(adlb, aes(x = ARMCD, y = AVAL)) +
  stat_summary(
    geom = "boxplot", fun.data = five_num,
    fun.args = list(probs = c(0.01, 0.25, 0.5, 0.75, 0.99)), fill = fc, color = oc
  ) +
  stat_summary(geom = "point", fun = mean, size = 3, shape = 8) +
  geom_point(data = adlb_o, aes(x = ARMCD, y = AVAL), shape = 1) +
  geom_text(data = adlb_o, aes(x = ARMCD, y = AVAL, label = ID), size = 3, hjust = -0.2) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    subtitle = paste("Visit:", adlb$AVISIT[1]),
    caption = "The whiskers extend to the largest and smallest observed value within 1.5*IQR.",
    x = "Treatment Group",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

plot <- bp_annos(bp_5, oc)
plot
```

`r webr_code_labels <- c("setup", "plot4")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot Specifying <br/> Marker for Mean

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot5, test = list(plot_v5 = "plot")}
bp_6 <- ggplot(adlb, aes(x = ARMCD, y = AVAL)) +
  geom_boxplot(fill = fc, color = oc) +
  stat_summary(geom = "point", fun = mean, size = 3, shape = 5) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    subtitle = paste("Visit:", adlb$AVISIT[1]),
    caption = "The whiskers extend to the minimum and maximum values.",
    x = "Treatment Group",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

plot <- bp_annos(bp_6, oc)
plot
```

`r webr_code_labels <- c("setup", "plot5")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot by Treatment <br/> and Timepoint

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->

```{r plot6, test = list(plot_v6 = "plot"), opts.label = "skip_test_strict"}
adsl <- random.cdisc.data::cadsl
adlb <- random.cdisc.data::cadlb

adlb_v <- adlb %>%
  filter(PARAMCD == "ALT" & AVISIT %in% c("WEEK 1 DAY 8", "WEEK 2 DAY 15", "WEEK 3 DAY 22", "WEEK 4 DAY 29"))

bp_7 <- ggplot(adlb_v, aes(x = AVISIT, y = AVAL)) +
  stat_summary(
    geom = "boxplot",
    fun.data = five_num,
    position = position_dodge2(.5),
    aes(fill = ARMCD, color = ARMCD)
  ) +
  stat_summary(
    geom = "point",
    fun = mean,
    aes(group = ARMCD),
    size = 3,
    shape = 8,
    position = position_dodge2(1)
  ) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    caption = "The whiskers extend to the minimum and maximum values.",
    x = "Visit",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

plot <- bp_annos(bp_7, oc, 2)
plot
```

`r webr_code_labels <- c("setup", "plot6")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot by Timepoint <br/> and Treatment

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->

```{r plot7, test = list(plot_v7 = "plot"), opts.label = "skip_test_strict"}
bp_8 <- ggplot(adlb_v, aes(x = ARMCD, y = AVAL)) +
  stat_summary(
    geom = "boxplot",
    fun.data = five_num,
    position = position_dodge2(width = .5),
    aes(fill = AVISIT, color = AVISIT)
  ) +
  stat_summary(
    geom = "point",
    fun = mean,
    aes(group = AVISIT),
    size = 3,
    shape = 8,
    position = position_dodge2(1)
  ) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    caption = "The whiskers extend to the minimum and maximum values.",
    x = "Treatment Group",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

plot <- bp_annos(bp_8, oc, 2)
plot
```

`r webr_code_labels <- c("setup", "plot7")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot with <br/> Table Section

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot8, test = list(plot_v8 = "plot")}
# Make wide dataset with summary statistics
adlb_wide <- adlb %>%
  group_by(ARMCD) %>%
  summarise(
    n = n(),
    mean = round(mean(AVAL), 1),
    median = round(median(AVAL), 1),
    median_ci = paste(round(DescTools::MedianCI(AVAL)[1:2], 1), collapse = " - "),
    q1_q3 = paste(round(quantile(AVAL, c(0.25, 0.75)), 1), collapse = " - "),
    min_max = paste0(round(min(AVAL), 1), "-", max(round(AVAL, 2)))
  )

# Make long dataset
adlb_long <- tidyr::gather(adlb_wide, key = type, value = stat, n:min_max)
adlb_long <- adlb_long %>% mutate(
  type_lbl = case_when(
    type == "n" ~ "n",
    type == "mean" ~ "Mean",
    type == "median" ~ "Median",
    type == "median_ci" ~ "95% CI for Median",
    type == "q1_q3" ~ "25% and 75%-ile",
    type == "min_max" ~ "Min - Max"
  )
)
adlb_long$type_lbl <- factor(adlb_long$type_lbl,
  levels = c("Min - Max", "25% and 75%-ile", "95% CI for Median", "Median", "Mean", "n")
)

bp_9 <- ggplot(adlb, aes(x = ARMCD, y = AVAL)) +
  stat_summary(geom = "boxplot", fun.data = five_num, fill = fc, color = oc) +
  stat_summary(geom = "point", fun = mean, size = 3, shape = 8) +
  labs(
    title = "Box Plot of Laboratory Test Results",
    subtitle = paste("Visit:", adlb$AVISIT[1]),
    x = "Treatment Group",
    y = paste0(adlb$PARAMCD[1], " (", adlb$AVALU[1], ")")
  ) +
  theme_bp

tbl_theme <- theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.ticks = element_blank(),
  axis.title = element_blank(),
  axis.text.y = element_text(face = "plain"),
  axis.text.x = element_blank()
)

tbl_1 <- ggplot(adlb_long, aes(x = ARMCD, y = type_lbl, label = stat), vjust = 10) +
  geom_text(size = 3) +
  scale_y_discrete(labels = levels(adlb_long$type_lbl)) +
  theme_bw() +
  tbl_theme +
  labs(caption = "The whiskers extend to the minimum and maximum values.") +
  theme_bp

plot <- cowplot::plot_grid(bp_annos(bp_9, oc), tbl_1,
  rel_heights = c(4, 2),
  ncol = 1, nrow = 2, align = "v"
)
plot
```

`r webr_code_labels <- c("setup", "plot8")` {{< include ../../_utils/webr.qmd >}}
:::

## Plot with Number of Patients <br/> only in Table Section

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot9, test = list(plot_v9 = "plot")}
tbl_2 <- adlb_long %>%
  filter(type == "n") %>%
  ggplot(aes(x = ARMCD, y = type_lbl, label = stat)) +
  geom_text(size = 3) +
  scale_y_discrete(labels = "n") +
  theme_bw() +
  tbl_theme +
  labs(caption = "The whiskers extend to the minimum and maximum values.") +
  theme_bp

plot <- cowplot::plot_grid(bp_annos(bp_9, oc), tbl_2,
  rel_heights = c(6, 1),
  ncol = 1, nrow = 2, align = "v"
)
plot
```

`r webr_code_labels <- c("setup", "plot8", "plot9")` {{< include ../../_utils/webr.qmd >}}
:::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

```{r teal, opts.label = c("skip_if_testing", "app"), eval = packageVersion("teal.modules.general") >= "0.3.0"}
library(teal.modules.general)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(tern)

  ADSL <- random.cdisc.data::cadsl
  ADLB <- random.cdisc.data::cadlb

  # If PARAMCD and AVISIT are not factors, convert to factors
  # Also fill in missing values with "<Missing>"
  ADLB <- ADLB %>%
    df_explicit_na(
      omit_columns = setdiff(names(ADLB), c("PARAMCD", "AVISIT")),
      char_as_factor = TRUE
    )

  # If statment below fails, pre-process ADLB to be one record per
  # study, subject, param and visit eg. filter with ANLFL = 'Y'
  stopifnot(nrow(ADLB) == nrow(unique(ADLB[, c("STUDYID", "USUBJID", "PARAMCD", "AVISIT")])))
})
datanames <- c("ADSL", "ADLB")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADSL <- data[["ADSL"]]
ADLB <- data[["ADLB"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_g_bivariate(
      x = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          label = "Select variable:",
          choices = names(ADSL),
          selected = "ARMCD",
          fixed = FALSE
        )
      ),
      y = data_extract_spec(
        dataname = "ADLB",
        filter = list(
          filter_spec(
            vars = c("PARAMCD"),
            choices = levels(ADLB$PARAMCD),
            selected = levels(ADLB$PARAMCD)[1],
            multiple = FALSE,
            label = "Choose lab parameter"
          ),
          filter_spec(
            vars = c("AVISIT"),
            choices = levels(ADLB$AVISIT),
            selected = levels(ADLB$AVISIT)[1],
            multiple = FALSE,
            label = "Choose visit"
          )
        ),
        select = select_spec(
          label = "Select variable:",
          choices = names(ADLB),
          selected = "AVAL",
          multiple = FALSE,
          fixed = TRUE
        )
      ),
      row_facet = data_extract_spec(
        dataname = "ADLB",
        select = select_spec(
          label = "Select variables:",
          choices = names(ADLB),
          selected = NULL,
          multiple = FALSE,
          fixed = FALSE
        )
      ),
      col_facet = data_extract_spec(
        dataname = "ADLB",
        select = select_spec(
          label = "Select variables:",
          choices = names(ADLB),
          selected = NULL,
          multiple = FALSE,
          fixed = FALSE
        )
      )
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../_utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}
:::
