---
title: IPPG01
subtitle: Individual Patient Plot Over Time
---

------------------------------------------------------------------------

{{< include ../../test-utils/envir_hook.qmd >}}

For illustration purposes, we will subset the `adlb` dataset for safety population in treatment arm A and a specific lab parameter (`ALT`).

::: panel-tabset
## Data Setup

```{r setup, message=FALSE}
#| code-fold: show

library(tern)
library(scda)
library(dplyr)
library(ggplot2)
library(nestcolor)

# use small sample size
adsl <- synthetic_cdisc_dataset("latest", "adsl") %>% slice(1:15)
adlb <- synthetic_cdisc_dataset("latest", "adlb") %>% filter(USUBJID %in% adsl$USUBJID)

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adlb <- df_explicit_na(adlb)

adlb_f <- adlb %>%
  filter(
    SAFFL == "Y",
    PARAMCD == "ALT",
    AVISIT != "SCREENING",
    ARMCD == "ARM A"
  ) %>%
  mutate(Patient_ID = sub(".*id-", "", USUBJID))
```

## Standard Plot

The user can select different `plotting_choices` depending on their preference.
To demonstrate, separate plots are produced with a maximum of 3 observations each.

```{r plots1, test = list(plots_v1 = "plots")}
plots <- g_ipp(
  df = adlb_f,
  xvar = "AVISIT",
  yvar = "AVAL",
  xlab = "Visit",
  ylab = "SGOT/ALT (U/L)",
  id_var = "Patient_ID",
  title = "Individual Patient Plots",
  subtitle = "Treatment Arm A",
  plotting_choices = "split_by_max_obs",
  max_obs_per_plot = 3
)
plots
```

## Plot with Patient Baselines as Reference

Here, patients' individual baseline values will be shown for reference.
Note that users can provide their own custom theme to the function via the `ggtheme` argument.

```{r plots2, test = list(plots_v2 = "plots")}
plots <- g_ipp(
  df = adlb_f,
  xvar = "AVISIT",
  yvar = "AVAL",
  xlab = "Visit",
  ylab = "SGOT/ALT (U/L)",
  id_var = "Patient_ID",
  title = "Individual Patient Plots",
  subtitle = "Treatment Arm A",
  add_baseline_hline = TRUE,
  yvar_baseline = "BASE",
  ggtheme = theme_minimal(),
  plotting_choices = "split_by_max_obs",
  max_obs_per_plot = 3
)
plots
```

{{< include ../../test-utils/save_results.qmd >}}

## `teal` App

```{r teal, opts.label=c('skip_if_testing', 'app')}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(tern)
  library(scda)
  library(dplyr)

  # use small sample size
  ADSL <- synthetic_cdisc_dataset("latest", "adsl") %>% slice(1:15)
  ADLB <- synthetic_cdisc_dataset("latest", "adlb") %>% filter(USUBJID %in% ADSL$USUBJID)

  # Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
  ADSL <- df_explicit_na(ADSL)
  ADLB <- df_explicit_na(ADLB) %>%
    filter(AVISIT != "SCREENING")
})
datanames <- c("ADSL", "ADLB")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADLB <- data[["ADLB"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_g_ipp(
      label = "Individual Patient Plot",
      dataname = "ADLB",
      arm_var = choices_selected(
        value_choices(ADLB, c("ARMCD")),
        "ARM A"
      ),
      paramcd = choices_selected(
        value_choices(ADLB, "PARAMCD"),
        "ALT"
      ),
      aval_var = choices_selected(
        variable_choices(ADLB, c("AVAL")),
        "AVAL"
      ),
      avalu_var = choices_selected(
        variable_choices(ADLB, c("AVALU")),
        "AVALU",
        fixed = TRUE
      ),
      id_var = choices_selected(
        variable_choices(ADLB, c("USUBJID")),
        "USUBJID",
        fixed = TRUE
      ),
      visit_var = choices_selected(
        variable_choices(ADLB, c("AVISIT")),
        "AVISIT"
      ),
      baseline_var = choices_selected(
        variable_choices(ADLB, c("BASE")),
        "BASE",
        fixed = TRUE
      ),
      add_baseline_hline = FALSE,
      separate_by_obs = FALSE
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../repro.qmd >}}
:::
