---
title: LTG01
subtitle: Lattice Plot of Laboratory Tests by Treatment Group Over Time
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

Lattice plots are natively handled by R, the examples below rely mostly on the package `ggplot2`.

:::: {.panel-tabset}
## Data Setup

```{r setup, message = FALSE}
#| code-fold: show

library(scda)
library(scda.2022)
library(tern)
library(teal.modules.clinical)
library(ggplot2)
library(dplyr)
library(nestcolor)

# Datasets
adsl <- synthetic_cdisc_dataset("latest", "adsl") %>% slice(1:8)
adlb <- synthetic_cdisc_dataset("latest", "adlb") %>% filter(USUBJID %in% adsl$USUBJID)

# Pre-processing
adlb$AVISIT_txt <- adlb$AVISIT
adlb$AVISIT <- as.numeric(adlb$AVISIT)
adlb$ARM_N <- adlb$ARM
levels(adlb$ARM_N) <- with(
  data = adlb,
  paste0(
    levels(ARM_N), " (N = ",
    tapply(SUBJID, ARM_N, function(x) length(unique(x))), ")"
  )
)

# Plot utils
npch <- 1:25
npatients <- length(unique(adlb$SUBJID))
pch <- c(
  rep(npch, times = npatients %/% length(npch)),
  npch[1:(npatients %% length(npch))]
)
```

## Plot of Liver Function Tests

#### Basic Plot

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->
```{r plot1, test = list(plot_v1 = "plot"), opts.label = "skip_test_strict"}
# General mapping and "lattice" ("facet" in ggplot2 nomenclature).
g1 <- {
  ggplot(
    data = adlb,
    mapping = aes(x = AVISIT, y = AVAL, colour = SUBJID, shape = SUBJID)
  ) +
    facet_grid(LBTESTCD ~ ARM, scales = "free_y") +
    scale_shape_manual(values = pch)
}

# Add points and lines.
g1 <- g1 + geom_point()
g1 <- g1 + geom_line()
plot <- g1
plot
```

`r webr_code_labels <- c("setup", "plot1")`
{{< include ../../_utils/webr.qmd >}}
:::

#### Modifying Facets

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

The units describing rows of panes and the number of patients under each arm is specified by modifying `facet_grid()`:

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->
```{r plot2, test = list(plot_v2 = "plot"), opts.label = "skip_test_strict"}
# Include the units and the sample size N.
g2 <- g1 + facet_grid(
  paste0(LBTESTCD, "\n(", AVALU, ")") ~ ARM_N,
  scales = "free_y"
)

plot <- g2
plot
```

`r webr_code_labels <- c("setup", "plot1", "plot2")`
{{< include ../../_utils/webr.qmd >}}
:::

#### Modifying X-Axis

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

The graphic elements are modified through usual `ggplot2` functions.
For instance, the x-axis could be improved as follows:

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->
```{r plot3, test = list(plot_v3 = "g3"), opts.label = "skip_test_strict"}
g3 <- g2 + theme(
  axis.text.x = element_text(angle = 45, hjust = 1),
  axis.title = element_blank()
) + scale_x_continuous(breaks = adlb$AVISIT, labels = adlb$AVISIT_txt)

plot <- g3
plot
```

`r webr_code_labels <- c("setup", "plot1", "plot2", "plot3")`
{{< include ../../_utils/webr.qmd >}}
:::

## Plot of Liver Function Tests <br/> Including Mean, Median, and 95% CIs

The functions `stat_mean_ci` and `stat_median_ci` from the `tern` package allow the addition of mean and/or median confidence intervals.
The example below suggests a larger dataset, where the individual subject legend may not be relevant but the mean or the median are of special interest.

#### Pre-Processing

```{r pre-processing}
#| code-fold: show

# Datasets
adsl <- synthetic_cdisc_dataset("latest", "adsl") %>% slice(1:40)
adlb <- synthetic_cdisc_dataset("latest", "adlb") %>% filter(USUBJID %in% adsl$USUBJID)

# Pre-processing
adlb$AVISIT_txt <- adlb$AVISIT
adlb$AVISIT <- as.numeric(adlb$AVISIT)
adlb$ARM_N <- adlb$ARM
levels(adlb$ARM_N) <- with(
  data = adlb,
  paste0(
    levels(ARM_N), " (N = ",
    tapply(SUBJID, ARM_N, function(x) length(unique(x))), ")"
  )
)

# Plot utils
npch <- 1:25
npatients <- length(unique(adlb$SUBJID))
pch <- c(
  rep(npch, times = npatients %/% length(npch)),
  npch[1:(npatients %% length(npch))]
)
```

#### Basic Plot

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot4, test = list(plot_v4 = "plot")}
# General mapping and "lattice" ("facet" in ggplot2 nomenclature)
g4 <- {
  ggplot(
    data = adlb,
    mapping = aes(x = AVISIT, y = AVAL, colour = SUBJID, shape = SUBJID)
  ) +
    facet_grid(LBTESTCD ~ ARM_N, scales = "free_y") +
    scale_shape_manual(values = pch) +
    scale_color_manual(values = rep(getOption("ggplot2.discrete.colour"), 2))
}

# Add points and lines.
# Note that with so many patients, legend might not be useful and transparency
#   is advisable.
g4 <- g4 + geom_point(alpha = .3)
g4 <- g4 + geom_line(alpha = .3)
g4 <- g4 + guides(colour = "none", shape = "none")
g4 <- g4 + theme(
  axis.text.x = element_text(angle = 45, hjust = 1),
  axis.title  = element_blank()
)
g4 <- g4 + scale_x_continuous(breaks = adlb$AVISIT, labels = adlb$AVISIT_txt)
plot <- g4
plot
```

`r webr_code_labels <- c("setup", "pre-processing", "plot4")`
{{< include ../../_utils/webr.qmd >}}
:::

#### Adding Mean

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->
```{r plot51, test = list(plot_v51 = "plot"), opts.label = "skip_test_strict"}
# Add the mean along with the 95% CI at every visit.
g51 <- g4 + stat_summary(
  fun = mean, linewidth = 1, geom = "line",
  aes(group = 1, linetype = "Mean +/- 95% CI")
)
g51 <- g51 + stat_summary(
  fun.data = tern::stat_mean_ci, geom = "errorbar",
  aes(group = 1, linetype = "Mean +/- 95% CI")
)
plot <- g51 + guides(linetype = guide_legend(title = NULL))
plot
```

`r webr_code_labels <- c("setup", "pre-processing", "plot4", "plot51")`
{{< include ../../_utils/webr.qmd >}}
:::

#### Adding Median

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->
```{r plot52, test = list(plot_v52 = "plot"), opts.label = "skip_test_strict"}
# Instead of a mean, the median could be more relevant.
g52 <- g51 + stat_summary(
  fun = median, linewidth = 1, geom = "line",
  aes(group = 1, linetype = "Median +/- 95% CI")
)
g52 <- g52 + stat_summary(
  fun.data = tern::stat_median_ci, geom = "errorbar",
  aes(group = 1, linetype = "Median +/- 95% CI")
)
plot <- g52 + guides(linetype = guide_legend(title = "Aggregate"))
plot
```

`r webr_code_labels <- c("setup", "pre-processing", "plot4", "plot51", "plot52")`
{{< include ../../_utils/webr.qmd >}}
:::

#### Changing Confidence Level

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->
```{r plot53, test = list(plot_v53 = "plot"), opts.label = "skip_test_strict"}
# Change the confidence level of interval for the median.
# Note: check `?stat_mean_ci()` and `?stat_median_ci()` for further fine tuning.
g53 <- g4 + stat_summary(
  fun = median, linewidth = 1, geom = "line",
  aes(group = 1, linetype = "Median +/- 80% CI")
)
g53 <- g53 + stat_summary(
  fun.data = function(x) tern::stat_median_ci(x, conf_level = 0.8),
  geom = "errorbar", aes(group = 1, linetype = "Median +/- 80% CI")
)
plot <- g53 + guides(linetype = guide_legend(title = NULL))
plot
```

`r webr_code_labels <- c("setup", "pre-processing", "plot4", "plot53")`
{{< include ../../_utils/webr.qmd >}}
:::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

```{r teal, opts.label = c("skip_if_testing", "app")}
#| code-fold: show

# In progress
```

{{< include ../../repro.qmd >}}

::::
