---
title: MNG01
subtitle: Mean Plot
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
library(dplyr)
library(tern)
library(nestcolor)

adsl <- random.cdisc.data::cadsl
adlb <- random.cdisc.data::cadlb
advs <- random.cdisc.data::cadvs

adsl_f <- adsl %>%
  filter(SAFFL == "Y") %>%
  df_explicit_na()

adlb_f <- adlb %>%
  mutate(AVISIT = forcats::fct_reorder(AVISIT, AVISITN, min)) %>%
  dplyr::filter(
    SAFFL == "Y",
    ANL01FL == "Y",
    ONTRTFL == "Y",
    PARAMCD == "ALT",
    AVISIT != "SCREENING"
  ) %>%
  droplevels() %>%
  df_explicit_na()

advs_f <- advs %>%
  mutate(AVISIT = forcats::fct_reorder(AVISIT, AVISITN, min)) %>%
  dplyr::filter(
    SAFFL == "Y",
    ONTRTFL == "Y",
    PARAMCD == "DIABP",
    AVISIT != "SCREENING"
  ) %>%
  droplevels() %>%
  df_explicit_na()
```

```{r include = FALSE}
webr_code_labels <- c("setup")
```

{{< include ../../_utils/webr_no_include.qmd >}}

## Output

:::::::::: panel-tabset
## Plot of Mean <br/> and CI

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot1, test = list(plot_v1 = "plot")}
plot <- g_lineplot(
  df = adlb_f,
  alt_counts_df = adsl_f,
  subtitle = "Laboratory Test:"
)
plot
```

```{r include = FALSE}
webr_code_labels <- c("plot1")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Plot of Mean and CIs of Change <br/> from Baseline (Changing the Input <br/> Analysis Data Set and Analysis Variable)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot2, test = list(plot_v2 = "plot")}
plot <- g_lineplot(
  df = advs_f,
  alt_counts_df = adsl_f,
  variables = control_lineplot_vars(y = "CHG"),
  title = "Plot of Mean and 95% Confidence Intervals of Change from Baseline by Visit",
  subtitle = "Assessment:"
)
plot
```

```{r include = FALSE}
webr_code_labels <- c("plot2")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Plot of Mean (+/-SD) <br/> (Changing the Statistics)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot3, test = list(plot_v3 = "plot")}
plot <- g_lineplot(
  df = adlb_f,
  alt_counts_df = adsl_f,
  interval = "mean_sdi",
  whiskers = c("mean_sdi_lwr", "mean_sdi_upr"),
  title = "Plot of Mean (+/-SD)",
  subtitle = "Laboratory Test:"
)
plot
```

```{r include = FALSE}
webr_code_labels <- c("plot3")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Plot of Mean and CI <br/> (Modify Alpha Level)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot4, test = list(plot_v4 = "plot")}
plot <- g_lineplot(
  df = adlb_f,
  alt_counts_df = adsl_f,
  control = control_analyze_vars(conf_level = 0.80),
  title = "Plot of Mean and 80% Confidence Limits by Visit",
  subtitle = "Laboratory Test:"
)
plot
```

```{r include = FALSE}
webr_code_labels <- c("plot4")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Plot of Mean and CI (with Number <br/> of Patients only in Table Section)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot5, test = list(plot_v5 = "plot")}
plot <- g_lineplot(
  df = adlb_f,
  alt_counts_df = adsl_f,
  table = "n",
  subtitle = "Laboratory Test:"
)
plot
```

```{r include = FALSE}
webr_code_labels <- c("plot5")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Plot of Mean and CI <br/> (with Table Section)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot6, fig.height = 8, test = list(plot_v6 = "plot")}
plot <- g_lineplot(
  df = adlb_f,
  alt_counts_df = adsl_f,
  table = c("n", "mean", "mean_ci"),
  subtitle = "Laboratory Test:"
)
plot
```

```{r include = FALSE}
webr_code_labels <- c("plot6")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Plot of Median and CI (Visits <br/> Condensed in Table Section)

This option is not supported.

## Plot of Mean and <br/> Upper Confidence Limit

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r plot7, test = list(plot_v7 = "plot")}
plot <- g_lineplot(
  df = adlb_f,
  alt_counts_df = adsl_f,
  whiskers = "mean_ci_upr",
  subtitle = "Laboratory Test:"
)
plot
```

```{r include = FALSE}
webr_code_labels <- c("plot7")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Data Setup

```{r setup}
#| code-fold: show
```
::::::::::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

```{r teal, opts.label = c("skip_if_testing", "app"), eval = packageVersion("teal.modules.clinical") > "0.9.1"}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(dplyr)
  library(forcats)

  ADSL <- random.cdisc.data::cadsl
  ADLB <- random.cdisc.data::cadlb %>%
    mutate(AVISIT = fct_reorder(AVISIT, AVISITN, min))
})
datanames <- c("ADSL", "ADLB")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADSL <- data[["ADSL"]]
ADLB <- data[["ADLB"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_g_lineplot(
      label = "LINE PLOT",
      dataname = "ADLB",
      strata = choices_selected(
        variable_choices(ADSL, c("ARM", "ARMCD", "ACTARMCD")),
        "ARM"
      ),
      y = choices_selected(
        variable_choices(ADLB, c("AVAL", "BASE", "CHG", "PCHG")),
        "AVAL"
      ),
      param = choices_selected(
        value_choices(ADLB, "PARAMCD", "PARAM"),
        "ALT"
      )
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../_utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}
