---
title: DMT01
subtitle: Demographics and Baseline Characteristics
---

------------------------------------------------------------------------

{{< include ../../test-utils/envir_hook.qmd >}}

:::: {.panel-tabset}
## Data Setup

```{r setup, message = FALSE}
#| code-fold: show

library(tern)
library(scda)
library(scda.2022)
library(dplyr)
library(tidyr)

adsl <- synthetic_cdisc_dataset("latest", "adsl")
advs <- synthetic_cdisc_dataset("latest", "advs")
adsub <- synthetic_cdisc_dataset("latest", "adsub")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
advs <- df_explicit_na(advs)
adsub <- df_explicit_na(adsub)

# Change description in variable SEX.
adsl <- adsl %>%
  mutate(
    SEX = factor(case_when(
      SEX == "M" ~ "Male",
      SEX == "F" ~ "Female",
      SEX == "U" ~ "Unknown",
      SEX == "UNDIFFERENTIATED" ~ "Undifferentiated"
    )),
    AGEGR1 = factor(
      case_when(
        between(AGE, 18, 40) ~ "18-40",
        between(AGE, 41, 64) ~ "41-64",
        AGE > 64 ~ ">=65"
      ),
      levels = c("18-40", "41-64", ">=65")
    ),
    BMRKR1_CAT = factor(
      case_when(
        BMRKR1 < 3.5 ~ "LOW",
        BMRKR1 >= 3.5 & BMRKR1 < 10 ~ "MEDIUM",
        BMRKR1 >= 10 ~ "HIGH"
      ),
      levels = c("LOW", "MEDIUM", "HIGH")
    )
  ) %>%
  var_relabel(
    BMRKR1_CAT = "Biomarker 1 Categories"
  )
# The developer needs to do pre-processing to add necessary variables based on ADVS to analysis dataset.
# Obtain SBP, DBP and weight.
get_param_advs <- function(pname, plabel) {
  ds <- advs %>%
    filter(PARAM == plabel & AVISIT == "BASELINE") %>%
    select(USUBJID, AVAL)

  colnames(ds) <- c("USUBJID", pname)

  ds
}
# The developer needs to do pre-processing to add necessary variables based on ADSUB to analysis dataset.
# Obtain baseline BMI (BBMISI).
get_param_adsub <- function(pname, plabel) {
  ds <- adsub %>%
    filter(PARAM == plabel) %>%
    select(USUBJID, AVAL)

  colnames(ds) <- c("USUBJID", pname)

  ds
}
adsl <- adsl %>%
  left_join(get_param_advs("SBP", "Systolic Blood Pressure"), by = "USUBJID") %>%
  left_join(get_param_advs("DBP", "Diastolic Blood Pressure"), by = "USUBJID") %>%
  left_join(get_param_advs("WGT", "Weight"), by = "USUBJID") %>%
  left_join(get_param_adsub("BBMISI", "Baseline BMI"), by = "USUBJID")
```

## Table with an Additional <br/> Study-Specific Continuous Variable

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant1, test = list(result_v1 = "result")}
vars <- c("AGE", "AGEGR1", "SEX", "ETHNIC", "RACE", "BMRKR1")
var_labels <- c(
  "Age (yr)",
  "Age Group",
  "Sex",
  "Ethnicity",
  "Race",
  "Continous Level Biomarker 1"
)

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ACTARM") %>%
  add_overall_col("All Patients") %>%
  analyze_vars(
    vars = vars,
    var_labels = var_labels
  ) %>%
  build_table(adsl)

result
```

`r webr_code_labels <- c("setup", "variant1")`
{{< include ../../webr-utils/webr.qmd >}}
:::

## Table with an Additional <br/> Study-Specific Categorical Variable

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant2, test = list(result_v2 = "result")}
vars <- c("AGE", "AGEGR1", "SEX", "ETHNIC", "RACE", "BMRKR1_CAT")
var_labels <- c(
  "Age (yr)",
  "Age Group",
  "Sex",
  "Ethnicity",
  "Race",
  "Biomarker 1 Categories"
)

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ACTARM") %>%
  analyze_vars(
    vars = vars,
    var_labels = var_labels
  ) %>%
  build_table(adsl)

result
```

`r webr_code_labels <- c("setup", "variant2")`
{{< include ../../webr-utils/webr.qmd >}}
:::

## Table with Subgrouping <br/> for Some Analyses

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant3, test = list(result_v3 = "result")}
split_fun <- drop_split_levels

result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ACTARM") %>%
  analyze_vars(
    vars = c("AGE", "SEX", "RACE"),
    var_labels = c("Age", "Sex", "Race")
  ) %>%
  split_rows_by("STRATA1",
    split_fun = split_fun
  ) %>%
  analyze_vars("BMRKR1") %>%
  build_table(adsl)

result
```

`r webr_code_labels <- c("setup", "variant3")`
{{< include ../../webr-utils/webr.qmd >}}
:::

## Table with Additional Vital <br/> Signs Baseline Values

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant4, test = list(result_v4 = "result")}
result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ACTARM") %>%
  analyze_vars(
    vars = c("AGE", "SEX", "RACE", "DBP", "SBP"),
    var_labels = c(
      "Age (yr)",
      "Sex",
      "Race",
      "Diastolic Blood Pressure",
      "Systolic Blood Pressure"
    )
  ) %>%
  build_table(adsl)

result
```

`r webr_code_labels <- c("setup", "variant4")`
{{< include ../../webr-utils/webr.qmd >}}
:::

## Table with Additional <br/> Values from ADSUB

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant5, test = list(result_v5 = "result")}
result <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ACTARM") %>%
  analyze_vars(
    vars = c("AGE", "SEX", "RACE", "BBMISI"),
    var_labels = c(
      "Age (yr)",
      "Sex",
      "Race",
      "Baseline BMI"
    )
  ) %>%
  build_table(adsl)

result
```

`r webr_code_labels <- c("setup", "variant5")`
{{< include ../../webr-utils/webr.qmd >}}
:::

{{< include ../../test-utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

```{r teal, opts.label = c("skip_if_testing", "app")}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(scda)
  library(scda.2022)

  ADSL <- synthetic_cdisc_dataset("latest", "adsl")

  # Include `EOSDY` and `DCSREAS` variables below because they contain missing data.
  stopifnot(
    any(is.na(ADSL$EOSDY)),
    any(is.na(ADSL$DCSREAS))
  )
})
datanames <- "ADSL"
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_t_summary(
      label = "Demographic Table",
      dataname = "ADSL",
      arm_var = choices_selected(c("ARM", "ARMCD"), "ARM"),
      summarize_vars = choices_selected(
        c("SEX", "RACE", "BMRKR2", "EOSDY", "DCSREAS"),
        c("SEX", "RACE")
      ),
      useNA = "ifany"
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../webr-utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}

::::
