---
title: RSPT01
subtitle: Best Overall Response
---

------------------------------------------------------------------------

::: panel-tabset
## Data Setup

```{r setup, message=FALSE}
#| code-fold: show

library(dplyr)
library(tern)
library(scda)

adsl <- synthetic_cdisc_dataset("rcd_2022_10_13", "adsl")
adrs <- synthetic_cdisc_dataset("rcd_2022_10_13", "adrs")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adrs <- df_explicit_na(adrs)

anl_adrs <- adrs %>%
  filter(PARAMCD == "INVET") %>%
  select(STUDYID, USUBJID, PARAMCD, AVISIT, AVALC)
anl_adsl <- adsl %>%
  select(STUDYID, USUBJID, ARM, STRATA1)
```

Data pre-processing is done to label the analysis value (`AVALC`) so as to improve the readability of the output table.
In addition, the *response* is binary defined (`is_rsp`, responder yes/no) depending on the analysis value.
Finally, for comparison purposes, the reference arm is specified.

```{r}
#| code-fold: show

# Pre-Processing

anl <- anl_adsl %>%
  left_join(anl_adrs, by = c("STUDYID", "USUBJID")) %>%
  mutate(rsp_lab = d_onco_rsp_label(AVALC)) %>%
  mutate(is_rsp = AVALC %in% c("CR", "PR")) %>%
  mutate(ARM = relevel(ARM, ref = "A: Drug X")) %>%
  var_relabel(ARM = "Description of Planned Arm")
```

## Standard Table

The tabulation layout is built in layers for the analysis of overall response and applied to the pre-processed dataset.

```{r}
lyt_01 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    table_names = "test_prop_diff"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    table_names = "est_or"
  ) %>%
  estimate_multinomial_response(var = "rsp_lab")

result <- build_table(lyt = lyt_01, df = anl)
result
```

## Table Selecting <br/> Sections to Display

Remove (or add) rows of results by removing/adding the corresponding layers from the layout.
For instance, the odds-ratio row is removed by simply removing the `estimate_odds_ratio` call:

```{r}
lyt_02 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    table_names = "test_prop"
  ) %>%
  estimate_multinomial_response(var = "rsp_lab")

result <- build_table(lyt = lyt_02, df = anl)
result
```

## Table Modifying Analysis Details like Type of <br/> Confidence Interval, Alpha Level, Test Statistic

The confidence level is controlled by the `conf_level` parameter to the estimation functions.
Similarly, the methods for tests and confidence interval can be modified (see `?estimate_proportion_diff`).

```{r}
conf_level <- 0.90
lyt_03 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    conf_level = conf_level,
    method = "clopper-pearson",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    conf_level = conf_level,
    method = "ha",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    method = "fisher",
    table_names = "test_prop"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    conf_level = conf_level,
    table_names = "est_or"
  ) %>%
  estimate_multinomial_response(
    var = "rsp_lab",
    conf_level = conf_level,
    method = "clopper-pearson"
  )

result <- build_table(lyt = lyt_03, df = anl)
result
```

## Table with <br/> Stratified Analysis

The stratified analysis section can be added by defining the analyses needed with `control_binary_comparison` for the argument `strat_analysis` and identifying the stratification variables to use.

```{r}
strata <- "STRATA1"
lyt_04 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    table_names = "test_prop"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    table_names = "est_or"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Stratified Analysis",
    method = "cmh",
    variables = list(strata = strata),
    table_names = "est_prop_diff_strat"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    method = "cmh",
    variables = list(strata = strata),
    table_names = "test_prop_strat"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    variables = list(strata = strata, arm = "ARM"),
    table_names = "est_or_strat"
  ) %>%
  estimate_multinomial_response(var = "rsp_lab")

result <- build_table(lyt = lyt_04, df = anl)
result
```

## Table Modifying the Definition <br/> of Overall Response

The definition of *responders* is realized during the pre-processing step.
The layout does not need to be modified and can be reused.

```{r}
anl <- anl_adsl %>%
  left_join(anl_adrs, by = c("STUDYID", "USUBJID")) %>%
  mutate(rsp_lab = d_onco_rsp_label(AVALC)) %>%
  mutate(is_rsp = AVALC == "CR") %>%
  mutate(ARM = relevel(ARM, ref = "A: Drug X")) %>%
  var_relabel(ARM = "Description of Planned Arm")

result <- build_table(lyt = lyt_01, df = anl)
result
```

## Table Defining New <br/> Sections to Display

Similarly to in the previous tab, redefinition or relabeling of the result is a pre-processing step and the original table layout can be reused.

```{r}
anl <- anl_adsl %>%
  left_join(anl_adrs, by = c("STUDYID", "USUBJID")) %>%
  mutate(rsp_lab = as.character(d_onco_rsp_label(AVALC))) %>%
  mutate(
    rsp_lab = case_when(
      rsp_lab == "Complete Response (CR)" ~ "No Progression",
      rsp_lab == "Partial Response (PR)" ~ "No Progression",
      rsp_lab == "Stable Disease (SD)" ~ "No Progression",
      TRUE ~ rsp_lab
    )
  ) %>%
  mutate(is_rsp = rsp_lab %in% "No Progression") %>%
  mutate(ARM = relevel(ARM, ref = "A: Drug X")) %>%
  var_relabel(ARM = "Description of Planned Arm")

result <- build_table(lyt = lyt_01, df = anl)
result
```

## Teal

```{r teal, message=FALSE}
#| screenshot.opts = list(delay = 20)

library(teal.modules.clinical)
library(scda)
library(dplyr)

ADSL <- synthetic_cdisc_dataset("rcd_2022_10_13", "adsl")
ADRS <- synthetic_cdisc_dataset("rcd_2022_10_13", "adrs")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
ADSL <- df_explicit_na(ADSL)
ADRS <- df_explicit_na(ADRS)

ADSL <- ADSL %>%
  mutate(Dum_ARM = factor(rep("Single ARM", nrow(.))))
ADRS <- ADRS %>%
  mutate(Dum_ARM = factor(rep("Single ARM", nrow(.))))

arm_ref_comp <- list(
  ACTARMCD = list(
    ref = "ARM B",
    comp = c("ARM A", "ARM C")
  ),
  ARM = list(
    ref = "B: Placebo",
    comp = c("A: Drug X", "C: Combination")
  )
)

app <- init(
  data = cdisc_data(
    cdisc_dataset(
      dataname = "ADSL",
      x = ADSL,
      code =
        'ADSL <- df_explicit_na(synthetic_cdisc_dataset("rcd_2022_10_13", "adsl")) %>%
           mutate(Dum_ARM = factor(rep("Single ARM", nrow(.))))'
    ),
    cdisc_dataset(
      dataname = "ADRS",
      x = ADRS,
      code =
        'ADRS <- df_explicit_na(synthetic_cdisc_dataset("rcd_2022_10_13", "adrs")) %>%
          mutate(Dum_ARM = factor(rep("Single ARM", nrow(.))))'
    ),
    check = TRUE
  ),
  modules = modules(
    tm_t_binary_outcome(
      label = "Responders",
      dataname = "ADRS",
      paramcd = choices_selected(
        choices = value_choices(ADRS, "PARAMCD", "PARAM"),
        selected = "BESRSPI"
      ),
      arm_var = choices_selected(
        choices = variable_choices(ADRS, c("ARM", "ARMCD", "ACTARMCD", "Dum_ARM")),
        selected = "ARM"
      ),
      arm_ref_comp = arm_ref_comp,
      strata_var = choices_selected(
        choices = variable_choices(ADRS, c("SEX", "BMRKR2")),
        select = NULL
      ),
      rsp_table = TRUE
    )
  )
)

shinyApp(app$ui, app$server)
```
:::
