---
title: MHT01
subtitle: Medical History
---

------------------------------------------------------------------------

{{< include ../../test-utils/envir_hook.qmd >}}

::: panel-tabset
## Data Setup

```{r setup, message=FALSE}
#| code-fold: show

library(tern)
library(dplyr)
library(scda)

adsl <- synthetic_cdisc_dataset("latest", "adsl")
admh <- synthetic_cdisc_dataset("latest", "admh")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
admh <- df_explicit_na(admh)

adsl_f <- adsl %>%
  filter(SAFFL == "Y") %>%
  select(USUBJID, ACTARM)

admh_f <- admh %>%
  filter(SAFFL == "Y" & MHBODSYS != "" & MHDECOD != "") %>%
  var_relabel(
    MHBODSYS = "MedDRA System Organ Class",
    MHDECOD = "MedDRA Preferred Term"
  )
```

## Standard Table

```{r variant1, test = list(result_v1 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c("Total number of patients with at least one condition", "Total number of conditions")
  ) %>%
  split_rows_by(
    var = "MHBODSYS",
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(admh_f$MHBODSYS)
  ) %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c("Total number of patients with at least one condition", "Total number of conditions"),
    show_labels = "hidden"
  ) %>%
  count_occurrences(vars = "MHDECOD") %>%
  append_varlabels(admh_f, "MHDECOD", indent = 1L)

result <- build_table(lyt, admh_f, alt_counts_df = adsl_f) %>%
  prune_table()

result
```

## Table of History <br/> Prior to Study

```{r variant2, test = list(result_v2 = "result")}
admh_f_prior <- admh_f %>%
  filter(ASTDY <= 0)

result <- build_table(lyt, admh_f_prior, alt_counts_df = adsl_f) %>%
  prune_table()

result
```

## Table with Total Number <br/> of Conditions Suppressed

```{r variant3, test = list(result_v3 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique"),
    .labels = c("Total number of patients with at least one condition")
  ) %>%
  split_rows_by(
    var = "MHBODSYS",
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(admh_f$MHBODSYS)
  ) %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique"),
    .labels = c("Total number of patients with at least one condition"),
    show_labels = "hidden"
  ) %>%
  count_occurrences(vars = "MHDECOD") %>%
  append_varlabels(admh_f, "MHDECOD", indent = 1L)

result <- build_table(lyt, admh_f, alt_counts_df = adsl) %>%
  prune_table()

result
```

## Table with Total Number of Conditions Per Body <br/> System After The Summary of Patients

Not currently supported by `rtables`.
Users, please notify us if this variation is important to you.

## Table Showing Additional <br/> "All Patients" Column

Conditions are further sorted by decreasing high level terms and low level terms.

```{r variant4, test = list(result_v4 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  add_overall_col("All Patients") %>%
  analyze_num_patients(
    "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(unique = "Total number of patients with at least one event", nonunique = "Total number of conditions")
  ) %>%
  split_rows_by(
    var = "MHBODSYS",
    split_fun = split_fun,
    child_labels = "visible",
    label_pos = "topleft",
    split_label = obj_label(admh_f$MHBODSYS)
  ) %>%
  summarize_num_patients(
    "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(unique = "Total number of patients with at least one event", nonunique = "Total number of conditions")
  ) %>%
  count_occurrences(vars = "MHDECOD", .indent_mods = -1L) %>%
  append_varlabels(admh_f, "MHDECOD", indent = 1L)

scorefun_hlt <- cont_n_allcols
scorefun_llt <- score_occurrences_cols(col_indices = nlevels(adsl_f$ACTARM) + 1)

result <- build_table(lyt, admh_f, alt_counts_df = adsl_f) %>%
  prune_table() %>%
  sort_at_path(path = c("MHBODSYS"), scorefun = scorefun_hlt) %>%
  sort_at_path(path = c("MHBODSYS", "*", "MHDECOD"), scorefun = scorefun_llt)

result
```

{{< include ../../test-utils/save_results.qmd >}}

## `teal` App

```{r teal, message=FALSE, opts.label=c('skip_if_testing', 'app')}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(scda)

  ADSL <- synthetic_cdisc_dataset("latest", "adsl") %>%
    filter(SAFFL == "Y")
  ADMH <- synthetic_cdisc_dataset("latest", "admh") %>%
    filter(SAFFL == "Y" & MHBODSYS != "" & MHDECOD != "")
})
datanames <- c("ADSL", "ADMH")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADMH <- data[["ADMH"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_t_events(
      label = "Medical History Table",
      dataname = "ADMH",
      arm_var = choices_selected(c("ARM", "ARMCD"), "ARM"),
      llt = choices_selected(
        choices = variable_choices(ADMH, c("MHTERM", "MHDECOD")),
        selected = c("MHDECOD")
      ),
      hlt = choices_selected(
        choices = variable_choices(ADMH, c("MHBODSYS", "MHSOC")),
        selected = "MHBODSYS"
      ),
      add_total = TRUE,
      event_type = "condition"
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../repro.qmd >}}
:::
