---
title: AET02
subtitle: Adverse Events
---

------------------------------------------------------------------------

::: panel-tabset
## Data Setup

```{r setup, message=FALSE}
#| code-fold: show

library(dplyr)
library(tern)
library(scda)

adsl <- synthetic_cdisc_dataset("rcd_2022_10_13", "adsl")
adae <- synthetic_cdisc_dataset("rcd_2022_10_13", "adae")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adae <- df_explicit_na(adae)

# Define the split function
split_fun <- drop_split_levels
```

## Standard Table

```{r aet02_01_1}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  add_overall_col(label = "All Patients") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(
    vars = "AEDECOD",
    .indent_mods = -1L
  ) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
)
```

The variable `result` corresponds to the adverse events table.
However, it includes many empty rows accounting for events which were not reported.
The table can be post-processed to prune empty rows and to sort rows, for example by occurrence.

```{r aet02_01_2}
result <- result %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_onecol(4)
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )
result
```

## Table with <br/> Event Totals

```{r aet02_02}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  add_overall_col(label = "All Patients") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = -1L) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L) %>%
  count_values(
    "STUDYID",
    values = as.character(unique(adae$STUDYID)),
    .stats = "count",
    .labels = c(count = "Total number of events"),
    .indent_mods = c(count = -1L)
  )

result <- build_table(
  lyt,
  df = adae,
  # col_count = c(table(adsl$ARM), sum(table(adsl$ARM)))
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  ) %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_onecol(4)
  )

result
```

## Table with <br/> High-Level Term

```{r aet02_03}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  split_rows_by(
    "AEHLT",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun,
    indent_mod = -1L,
    label_pos = "topleft",
    split_label = obj_label(adae$AEHLT)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = c(count_fraction = -1L)) %>%
  append_varlabels(adae, c("AEDECOD"), indent = 2L)

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEHLT"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEHLT", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

result
```

## Table with Preferred <br/> Terms Only

```{r aet02_04}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = c(count_fraction = -1L)) %>%
  append_varlabels(adae, "AEDECOD")

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  sort_at_path(
    path = c("AEDECOD"), scorefun = score_occurrences
  )

result
```

## Table with Fill-In of <br/> Treatment Groups

```{r aet02_05}
adae_5 <- adae %>% dplyr::filter(ARM != "C: Combination")

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = c(count_fraction = -1L)) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt,
  df = adae_5,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

result
```

## Table of AEs with an Incidence Rate <br/> $\geq$ 5% in Any Treatment Group (subsetting <br/> preferred terms based on frequency)

```{r aet02_06}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    indent_mod = -1L,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  count_occurrences(
    vars = "AEDECOD",
    .indent_mods = c(count_fraction = 1L)
  ) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

criteria_fun <- function(tr) is(tr, "ContentRow")
result <- trim_rows(result, criteria = criteria_fun)

row_condition <- has_fraction_in_any_col(
  atleast = 0.05,
  col_names = names(table(adsl$ARM))
)
result <- prune_table(result, keep_rows(row_condition))

result
```

## Table of AEs with an Incidence Rate $\geq$ 5% in <br/> Any Treatment Group (subsetting preferred terms <br/> based on frequency with high-level terms)

```{r aet02_07}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  split_rows_by(
    "AEHLT",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEHLT)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  count_occurrences(vars = "AEDECOD") %>%
  append_varlabels(adae, c("AEDECOD"), indent = 2L)

result <- build_table(lyt, adae, alt_counts_df = adsl) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEHLT"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEHLT", "*", "AEDECOD"),
    scorefun = score_occurrences,
    decreasing = TRUE
  )

criteria_fun <- function(tr) is(tr, "ContentRow")
result <- trim_rows(result, criteria = criteria_fun)

row_condition <- has_fraction_in_any_col(
  atleast = 0.05,
  col_names = names(table(adsl$ARM))
)
result <- prune_table(result, keep_rows(row_condition))

result
```

## Table of AEs with an Incidence Rate $\geq$ 10% <br/> in Any Treatment Group (subsetting <br/> preferred terms based on frequency)

```{r aet02_08}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    indent_mod = -1L,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  count_occurrences(
    vars = "AEDECOD",
    .indent_mods = c(count_fraction = 1L)
  ) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

criteria_fun <- function(tr) is(tr, "ContentRow")
result <- trim_rows(result, criteria = criteria_fun)

row_condition <- has_fraction_in_any_col(
  atleast = 0.10,
  col_names = names(table(adsl$ARM))
)
result <- prune_table(result, keep_rows(row_condition))

result
```

## Table of AEs with an Incidence Rate $\geq$ 3 <br/> Patients in Any Treatment Group (subsetting <br/> preferred terms based on number of patients)

```{r aet02_09}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    indent_mod = -1L,
    split_fun = split_fun,
    split_label = obj_label(adae$AEBODSYS),
    label_pos = "topleft"
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  count_occurrences(
    vars = "AEDECOD",
    .indent_mods = c(count_fraction = 1L)
  ) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

criteria_fun <- function(tr) is(tr, "ContentRow")
result <- trim_rows(result, criteria = criteria_fun)

row_condition <- has_count_in_any_col(
  atleast = 3,
  col_names = names(table(adsl$ARM))
)
result <- prune_table(result, keep_rows(row_condition))

result
```

## Table of AEs with a Difference in Incidence Rate $\geq$ 5% <br/> Between Any Treatment (subsetting preferred terms based <br/> on difference in percentage between treatment groups)

```{r aet02_10}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    indent_mod = -1L,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  count_occurrences(
    vars = "AEDECOD",
    .indent_mods = c(count_fraction = 1L)
  ) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

criteria_fun <- function(tr) is(tr, "ContentRow")
result <- trim_rows(result, criteria = criteria_fun)

row_condition <- has_fractions_difference(
  atleast = 0.05,
  col_names = levels(adsl$ARM)
)
result <- prune_table(result, keep_rows(row_condition))
result
```

## Table of AEs with an Incidence Rate $\geq$ 5% <br/> in B: Placebo (subsetting preferred terms based <br/> on frequency for a particular treatment group)

```{r aet02_11}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    indent_mod = -1L,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  count_occurrences(
    vars = "AEDECOD",
    .indent_mods = c(count_fraction = 1L)
  ) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)
result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )
criteria_fun <- function(tr) is(tr, "ContentRow")
result <- trim_rows(result, criteria = criteria_fun)
row_condition <- has_fraction_in_cols(
  atleast = 0.05,
  col_names = c("B: Placebo")
)
result <- prune_table(result, keep_rows(row_condition))
result
```

## Table of AEs with a Difference in Incidence Rate $\geq$ 5% Between <br/> Arm A and Arm B or Arm C (displaying preferred terms with a <br/> difference of at least x% between selected treatment groups)

```{r aet02_12}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    indent_mod = -1L,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "Total number of patients with at least one adverse event"
  ) %>%
  count_occurrences(
    vars = "AEDECOD",
    .indent_mods = c(count_fraction = 1L)
  ) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

criteria_fun <- function(tr) is(tr, "ContentRow")
result <- trim_rows(result, criteria = criteria_fun)

row_condition1 <- has_fractions_difference(atleast = 0.05, col_names = c("A: Drug X", "B: Placebo"))
row_condition2 <- has_fractions_difference(atleast = 0.05, col_names = c("A: Drug X", "C: Combination"))
row_condition <- row_condition1 | row_condition2

result <- prune_table(result, keep_rows(row_condition))
result
```

## Teal

```{r teal, message=FALSE}
#| screenshot.opts = list(delay = 12)

library(teal.modules.clinical)
library(dplyr)
library(scda)

adsl <- synthetic_cdisc_dataset("rcd_2022_10_13", "adsl")
adae <- synthetic_cdisc_dataset("rcd_2022_10_13", "adae")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adae <- df_explicit_na(adae)

app <- teal::init(
  data = cdisc_data(
    cdisc_dataset("ADSL", adsl, code = 'ADSL <- df_explicit_na(synthetic_cdisc_dataset("rcd_2022_10_13", "adsl"))'),
    cdisc_dataset("ADAE", adae, code = 'ADAE <- df_explicit_na(synthetic_cdisc_dataset("rcd_2022_10_13", "adae"))'),
    check = TRUE
  ),
  modules = modules(
    tm_t_events(
      label = "Adverse Event Table",
      dataname = "ADAE",
      arm_var = choices_selected(c("ARM", "ARMCD"), "ARM"),
      llt = choices_selected(
        choices = variable_choices(adae, c("AETERM", "AEDECOD")),
        selected = c("AEDECOD")
      ),
      hlt = choices_selected(
        choices = variable_choices(adae, c("AEBODSYS", "AESOC")),
        selected = "AEBODSYS"
      ),
      add_total = TRUE,
      event_type = "adverse event"
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../si.qmd >}}

:::
