---
title: ENTXX
subtitle: Enrollment Variants
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
library(dplyr)
library(forcats)
library(tern)

adsl <- random.cdisc.data::cadsl

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)

adsl <- adsl %>%
  mutate(
    REGION1 = droplevels(REGION1),
    COUNTRY = droplevels(COUNTRY),
    INVID_INVNAM = paste(INVID, INVNAM, sep = " / ")
  ) %>%
  mutate(
    INVID = factor(INVID),
    SITEID = factor(SITEID),
    INVNAM = factor(INVNAM),
    INVID_INVNAM = factor(INVID_INVNAM)
  ) %>%
  var_relabel(
    INVID_INVNAM = "Investigator Number/Name",
    REGION1 = "Geographic Region 1",
    COUNTRY = "Country",
    INVID = "Investigator Identifier"
  ) %>%
  arrange(REGION1, COUNTRY, INVID)
```

```{r include = FALSE}
webr_code_labels <- c("setup")
```

{{< include ../../_utils/webr_no_include.qmd >}}

## Output

::::::: panel-tabset
## ENT01 -- Enrollment by Region, <br/> Country, and Investigator Number

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant1, test = list(result_v1 = "result")}
# Define the split function
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("REGION1", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$REGION1)) %>%
  summarize_row_groups() %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID", 2L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

```{r include = FALSE}
webr_code_labels <- c("variant1")
```

{{< include ../../_utils/webr.qmd >}}
:::

## ENT01A -- Enrollment by Country <br/> and Investigator Number

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant2, test = list(result_v2 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID", 1L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

```{r include = FALSE}
webr_code_labels <- c("variant2")
```

{{< include ../../_utils/webr.qmd >}}
:::

## ENT02 -- Enrollment by Region, Country, <br/> and Investigator Number/Name

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant3, test = list(result_v3 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("REGION1", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$REGION1)) %>%
  summarize_row_groups() %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID_INVNAM", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID_INVNAM", 2L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

```{r include = FALSE}
webr_code_labels <- c("variant3")
```

{{< include ../../_utils/webr.qmd >}}
:::

## ENT02A -- Enrollment by Country <br/> and Investigator Number/Name

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant4, test = list(result_v4 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID_INVNAM", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID_INVNAM", 1L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

```{r include = FALSE}
webr_code_labels <- c("variant4")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Data Setup

```{r setup}
#| code-fold: show
```
:::::::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

Note that for this module application, only the variables passed into `by_vars` are used when `row_groups` is selected. Variables passed into `analyze_vars` are additionally used when `row_groups` is deselected.

```{r teal, opts.label = c("skip_if_testing", "app")}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(dplyr)

  ADSL <- random.cdisc.data::cadsl

  # Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
  ADSL <- df_explicit_na(ADSL)

  adsl_labels <- col_labels(ADSL)

  ADSL <- ADSL %>%
    mutate(
      REGION1 = droplevels(REGION1),
      COUNTRY = droplevels(COUNTRY),
      INVID_INVNAM = paste(INVID, INVNAM, sep = " / ")
    ) %>%
    arrange(REGION1, COUNTRY, INVID) %>%
    mutate(
      INVID = factor(INVID),
      SITEID = factor(SITEID),
      INVNAM = factor(INVNAM),
      INVID_INVNAM = factor(INVID_INVNAM)
    )

  col_labels(ADSL) <- c(adsl_labels, c(INVID_INVNAM = "Investigator Number/Name"))
})
datanames <- "ADSL"
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADSL <- data[["ADSL"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_t_summary_by(
      label = "Enrollment Table",
      dataname = "ADSL",
      arm_var = choices_selected(
        choices = variable_choices(ADSL, c("ARM", "ARMCD")),
        selected = "ARM"
      ),
      by_vars = choices_selected(
        choices = variable_choices(ADSL, c("REGION1", "COUNTRY", "SITEID", "INVID", "INVNAM", "INVID_INVNAM")),
        selected = c("REGION1")
      ),
      summarize_vars = choices_selected(
        choices = variable_choices(ADSL, c("REGION1", "COUNTRY", "SITEID", "INVID", "INVNAM", "INVID_INVNAM")),
        selected = c("SITEID")
      ),
      denominator = choices_selected(
        choices = c("n", "N", "omit"),
        selected = "N",
        fixed = TRUE
      ),
      useNA = "ifany",
      row_groups = TRUE
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../_utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}
