---
title: ENTXX
subtitle: Enrollment Variants
---

------------------------------------------------------------------------

{{< include ../../test-utils/envir_hook.qmd >}}

::: panel-tabset
## Data Setup

```{r setup, message=FALSE}
#| code-fold: show

library(dplyr)
library(forcats)
library(scda)
library(tern)

adsl <- synthetic_cdisc_dataset("latest", "adsl")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)

adsl <- adsl %>%
  mutate(
    REGION1 = droplevels(REGION1),
    COUNTRY = droplevels(COUNTRY),
    INVID_INVNAM = paste(INVID, INVNAM, sep = " / ")
  ) %>%
  mutate(
    INVID = factor(INVID),
    SITEID = factor(SITEID),
    INVNAM = factor(INVNAM),
    INVID_INVNAM = factor(INVID_INVNAM)
  ) %>%
  var_relabel(
    INVID_INVNAM = "Investigator Number/Name",
    REGION1 = "Geographic Region 1",
    COUNTRY = "Country",
    INVID = "Investigator Identifier"
  ) %>%
  arrange(REGION1, COUNTRY, INVID)
```

## ENT01 -- Enrollment by Region, <br/> Country, and Investigator Number

```{r variant1, test = list(result_v1 = "result")}
# Define the split function
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("REGION1", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$REGION1)) %>%
  summarize_row_groups() %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID", 2L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

## ENT01A -- Enrollment by Country <br/> and Investigator Number

```{r variant2, test = list(result_v2 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID", 1L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

## ENT02 -- Enrollment by Region, Country, <br/> and Investigator Number/Name

```{r variant3, test = list(result_v3 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("REGION1", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$REGION1)) %>%
  summarize_row_groups() %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID_INVNAM", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID_INVNAM", 2L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

## ENT02A -- Enrollment by Country <br/> and Investigator Number/Name

```{r variant4, test = list(result_v4 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  split_rows_by("COUNTRY", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adsl$COUNTRY)) %>%
  summarize_row_groups() %>%
  analyze_vars("INVID_INVNAM", .stats = "count_fraction") %>%
  append_varlabels(adsl, "INVID_INVNAM", 1L)

result <- build_table(lyt, adsl) %>%
  prune_table()

result
```

{{< include ../../test-utils/save_results.qmd >}}

## `teal` App

Note that for this module application, only the variables passed into `by_vars` are used when `row_groups` is selected.
Variables passed into `analyze_vars` are additionally used when `row_groups` is deselected.

```{r teal, message=FALSE, opts.label='skip_if_testing'}
#| screenshot.opts = list(delay = 8)

library(teal.modules.clinical)

data <- teal_data()
data <- within(data, {
  library(dplyr)
  library(forcats)
  library(scda)
  ADSL <- synthetic_cdisc_dataset("latest", "adsl")

  # Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
  ADSL <- df_explicit_na(ADSL)

  adsl_labels <- var_labels(ADSL)

  ADSL <- ADSL %>%
    mutate(
      REGION1 = droplevels(REGION1),
      COUNTRY = droplevels(COUNTRY),
      INVID_INVNAM = paste(INVID, INVNAM, sep = " / ")
    ) %>%
    arrange(REGION1, COUNTRY, INVID) %>%
    mutate(
      INVID = factor(INVID),
      SITEID = factor(SITEID),
      INVNAM = factor(INVNAM),
      INVID_INVNAM = factor(INVID_INVNAM)
    )

  var_labels(ADSL) <- c(adsl_labels, "Investigator Number/Name")
})
datanames <- "ADSL"
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADSL <- data[["ADSL"]]

app <- init(
  data = data,
  modules = modules(
    tm_t_summary_by(
      label = "Enrollment Table",
      dataname = "ADSL",
      arm_var = choices_selected(
        choices = variable_choices(ADSL, c("ARM", "ARMCD")),
        selected = "ARM"
      ),
      by_vars = choices_selected(
        choices = variable_choices(ADSL, c("REGION1", "COUNTRY", "SITEID", "INVID", "INVNAM", "INVID_INVNAM")),
        selected = c("REGION1")
      ),
      summarize_vars = choices_selected(
        choices = variable_choices(ADSL, c("REGION1", "COUNTRY", "SITEID", "INVID", "INVNAM", "INVID_INVNAM")),
        selected = c("SITEID")
      ),
      denominator = choices_selected(
        choices = c("n", "N", "omit"),
        selected = "N",
        fixed = TRUE
      ),
      useNA = "ifany",
      row_groups = TRUE
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../repro.qmd >}}
:::
