---
title: LGRT02
subtitle: Multi-Variable Logistic Regression
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

::: panel-tabset
## Data Setup

```{r setup, message = FALSE}
#| code-fold: show

library(dplyr)
library(tern)

adsl <- random.cdisc.data::cadsl
adrs <- random.cdisc.data::cadrs

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adrs <- df_explicit_na(adrs)

adsl <- adsl %>%
  dplyr::filter(SEX %in% c("F", "M"))

adrs <- adrs %>%
  dplyr::filter(PARAMCD == "BESRSPI") %>%
  dplyr::mutate(
    Response = case_when(AVALC %in% c("PR", "CR") ~ 1, TRUE ~ 0),
    SEX = factor(SEX, c("M", "F")),
    RACE = factor(
      RACE,
      levels = c(
        "AMERICAN INDIAN OR ALASKA NATIVE", "ASIAN", "BLACK OR AFRICAN AMERICAN",
        "WHITE", "MULTIPLE", "NATIVE HAWAIIAN OR OTHER PACIFIC ISLANDER"
      )
    )
  ) %>%
  var_relabel(Response = "Response", SEX = "Sex", RACE = "Race")
```

## Multi-Variable Logistic Regression

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant1, test = list(result_v1 = "result")}
model <- fit_logistic(
  adrs,
  variables = list(response = "Response", arm = "ARMCD", covariates = c("SEX", "AGE"))
)
conf_level <- 0.95
df <- broom::tidy(model, conf_level = conf_level)

# empty string flag
df <- df_explicit_na(df, na_level = "_NA_")

result <- basic_table() %>%
  summarize_logistic(
    conf_level = conf_level,
    drop_and_remove_str = "_NA_"
  ) %>%
  append_topleft("Logistic regression") %>%
  build_table(df = df)
result
```

`r webr_code_labels <- c("setup", "variant1")` {{< include ../../_utils/webr.qmd >}}
:::

## Multi-Variable Logistic Regression <br/> with Interaction Term

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant2, test = list(result_v2 = "result")}
model <- fit_logistic(
  adrs,
  variables = list(
    response = "Response",
    arm = "ARMCD",
    covariates = c("SEX", "AGE"),
    interaction = "SEX"
  )
)
conf_level <- 0.95
df <- broom::tidy(model, conf_level = conf_level)

# empty string flag
df <- df_explicit_na(df, na_level = "_NA_")

result <- basic_table() %>%
  summarize_logistic(
    conf_level = conf_level,
    drop_and_remove_str = "_NA_"
  ) %>%
  append_topleft("Logistic regression with interaction") %>%
  build_table(df = df)
result
```

`r webr_code_labels <- c("setup", "variant2")` {{< include ../../_utils/webr.qmd >}}
:::

## Multi-Variable Logistic Regression <br/> Specifying Covariates

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant3, test = list(result_v3 = "result")}
model <- fit_logistic(
  adrs,
  variables = list(
    response = "Response",
    arm = "ARMCD",
    covariates = c("SEX", "AGE", "RACE")
  )
)
conf_level <- 0.95
df <- broom::tidy(model, conf_level = conf_level)

# empty string flag
df <- df_explicit_na(df, na_level = "_NA_")

result <- basic_table() %>%
  summarize_logistic(
    conf_level = conf_level,
    drop_and_remove_str = "_NA_"
  ) %>%
  append_topleft("y ~ ARM + SEX + AGE + RACE") %>%
  build_table(df = df)
result
```

`r webr_code_labels <- c("setup", "variant3")` {{< include ../../_utils/webr.qmd >}}
:::

## Multi-Variable Logistic Regression Setting <br/> an Event, Alpha Level, and Level for Interaction

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant4, test = list(result_v4 = "result")}
model <- fit_logistic(
  adrs,
  variables = list(
    response = "Response",
    arm = "ARMCD",
    covariates = c("SEX", "AGE"),
    interaction = "AGE"
  ),
  response_definition = "1 - response"
)
conf_level <- 0.9
df <- broom::tidy(model, conf_level = conf_level, at = c(30, 50))

# empty string flag
df <- df_explicit_na(df, na_level = "_NA_")

result <- basic_table() %>%
  summarize_logistic(
    conf_level = conf_level,
    drop_and_remove_str = "_NA_"
  ) %>%
  append_topleft("Estimations at age 30 and 50") %>%
  build_table(df = df)
result
```

`r webr_code_labels <- c("setup", "variant4")` {{< include ../../_utils/webr.qmd >}}
:::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

```{r teal, opts.label = c("skip_if_testing", "app")}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
    library(dplyr)

  ADSL <- random.cdisc.data::cadsl
  ADRS <- random.cdisc.data::cadrs %>%
    filter(PARAMCD %in% c("BESRSPI", "INVET"))
})
datanames <- c("ADSL", "ADRS")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADRS <- data[["ADRS"]]
arm_ref_comp <- list(
  ACTARMCD = list(
    ref = "ARM B",
    comp = c("ARM A", "ARM C")
  ),
  ARM = list(
    ref = "B: Placebo",
    comp = c("A: Drug X", "C: Combination")
  )
)

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_t_logistic(
      label = "Logistic Regression",
      dataname = "ADRS",
      arm_var = choices_selected(
        choices = variable_choices(ADRS, c("ARM", "ARMCD")),
        selected = "ARM"
      ),
      arm_ref_comp = arm_ref_comp,
      paramcd = choices_selected(
        choices = value_choices(ADRS, "PARAMCD", "PARAM"),
        selected = "BESRSPI"
      ),
      cov_var = choices_selected(
        choices = c("SEX", "AGE", "BMRKR1", "BMRKR2"),
        selected = "SEX"
      )
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../_utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}
:::
