---
title: ONCT05
subtitle: Objective Response Rate by Subgroup
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

::: panel-tabset
## Data Setup

```{r setup, message = FALSE}
#| code-fold: show

library(tern)
library(dplyr)

adsl <- random.cdisc.data::cadsl
adrs <- random.cdisc.data::cadrs

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adrs <- df_explicit_na(adrs)

adsl <- adsl %>%
  select(STUDYID, USUBJID, ARM, SEX, RACE, STRATA1, STRATA2)

adrs <- adrs %>%
  filter(PARAMCD == "INVET") %>%
  select(STUDYID, USUBJID, PARAMCD, AVALC)

anl <- inner_join(adsl, adrs, by = c("STUDYID", "USUBJID"))
anl_labels <- var_labels(anl)

anl <- anl %>%
  filter(ARM %in% c("A: Drug X", "B: Placebo")) %>%
  mutate(
    # Reorder levels of factor to make the placebo group the reference arm.
    ARM = relevel(ARM, ref = "B: Placebo") %>%
      droplevels()
  ) %>%
  droplevels() %>%
  mutate(rsp = AVALC %in% c("CR", "PR"))

var_labels(anl) <- c(anl_labels, rsp = "Is Response")
```

## Standard Table

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/insightsengineering/rtables/issues/830 -->

```{r variant1, test = list(result_v1 = "result"), opts.label = ifelse(packageVersion("rtables") < "0.6.6.9011", "skip_test_strict", "")}
df <- extract_rsp_subgroups(
  variables = list(
    rsp = "rsp",
    arm = "ARM",
    subgroups = c("SEX", "STRATA2")
  ),
  data = anl
)

result <- basic_table() %>%
  tabulate_rsp_subgroups(df, vars = c("n_tot", "n", "n_rsp", "prop", "or", "ci"))

result
```

`r webr_code_labels <- c("setup", "variant1")` {{< include ../../_utils/webr.qmd >}}
:::

## Table Specifying <br/> Class Variables

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

Here, the levels of subgroup variables `SEX` and `STRATA1` are reordered.
`STRATA1` is reordered by frequency.

<!-- skip strict because of https://github.com/insightsengineering/rtables/issues/830 -->

```{r variant2, test = list(result_v2 = "result"), opts.label = ifelse(packageVersion("rtables") < "0.6.6.9011", "skip_test_strict", "")}
anl_reorder <- anl %>%
  mutate(
    SEX = forcats::fct_relevel(SEX, "M", "F"),
    STRATA1 = forcats::fct_infreq(STRATA1)
  )

df <- extract_rsp_subgroups(
  variables = list(rsp = "rsp", arm = "ARM", subgroups = c("SEX", "STRATA1")),
  data = anl_reorder
)

result <- basic_table() %>%
  tabulate_rsp_subgroups(df, vars = c("n_tot", "n", "n_rsp", "prop", "or", "ci"))

result
```

`r webr_code_labels <- c("setup", "variant2")` {{< include ../../_utils/webr.qmd >}}
:::

## Table Selecting Columns <br/> and Changing the Alpha Level

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

<!-- skip strict because of https://github.com/insightsengineering/rtables/issues/830 -->

```{r variant3, test = list(result_v3 = "result"), opts.label = ifelse(packageVersion("rtables") < "0.6.6.9011", "skip_test_strict", "")}
df <- extract_rsp_subgroups(
  variables = list(
    rsp = "rsp",
    arm = "ARM",
    subgroups = c("SEX", "STRATA2")
  ),
  data = anl,
  conf_level = 0.9,
  method = "chisq"
)

result <- basic_table() %>%
  tabulate_rsp_subgroups(df, vars = c("n_tot", "n", "n_rsp", "prop", "or", "ci", "pval"))

result
```

`r webr_code_labels <- c("setup", "variant3")` {{< include ../../_utils/webr.qmd >}}
:::

## Table Setting Values <br/> Indicating Response

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

Create a new variable `new_rsp` in `anl` data that uses new criteria for responder.

<!-- skip strict because of https://github.com/insightsengineering/rtables/issues/830 -->

```{r variant4, test = list(result_v4 = "result"), opts.label = ifelse(packageVersion("rtables") < "0.6.6.9011", "skip_test_strict", "")}
anl_new <- anl %>%
  mutate(new_rsp = AVALC == "CR")

df <- extract_rsp_subgroups(
  variables = list(
    rsp = "new_rsp",
    arm = "ARM",
    subgroups = c("SEX", "STRATA2")
  ),
  data = anl_new
)

result <- basic_table() %>%
  tabulate_rsp_subgroups(df, vars = c("n_tot", "n", "n_rsp", "prop", "or", "ci"))

result
```

`r webr_code_labels <- c("setup", "variant4")` {{< include ../../_utils/webr.qmd >}}
:::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

```{r teal, opts.label = c("skip_if_testing", "app")}
# Use table, embedded in response forest plot module.
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  ADSL <- random.cdisc.data::cadsl
  ADRS <- random.cdisc.data::cadrs

  # Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
  ADSL <- df_explicit_na(ADSL)
  ADRS <- df_explicit_na(ADRS)
})
datanames <- c("ADSL", "ADRS")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADSL <- data[["ADSL"]]
ADRS <- data[["ADRS"]]
arm_ref_comp <- list(
  ARM = list(
    ref = "B: Placebo",
    comp = c("A: Drug X", "C: Combination")
  ),
  ARMCD = list(
    ref = "ARM B",
    comp = c("ARM A", "ARM C")
  )
)

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_g_forest_rsp(
      label = "Forest Response",
      dataname = "ADRS",
      arm_var = choices_selected(
        variable_choices(ADSL, c("ARM", "ARMCD")),
        "ARMCD"
      ),
      arm_ref_comp = arm_ref_comp,
      paramcd = choices_selected(
        value_choices(ADRS, "PARAMCD", "PARAM"),
        "BESRSPI"
      ),
      subgroup_var = choices_selected(
        variable_choices(ADSL, names(ADSL)),
        c("SEX")
      ),
      strata_var = choices_selected(
        variable_choices(ADSL, c("STRATA1", "STRATA2")),
        "STRATA2"
      ),
      plot_height = c(600L, 200L, 2000L)
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../_utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}
:::
