---
title: RBMIT01
subtitle: Tables for RBMI
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

:::: {.panel-tabset}

```{r setup, echo = FALSE}
library(tern.rbmi)
library(dplyr)
set.seed(123)

data <- antidepressant_data
levels(data$THERAPY) <- c("PLACEBO", "DRUG") # This is important! The order defines the computation order later

missing_var <- "CHANGE"
vars <- list(
  id = "PATIENT",
  visit = "VISIT",
  expand_vars = c("BASVAL", "THERAPY"),
  group = "THERAPY"
)
covariates <- list(
  draws = c("BASVAL*VISIT", "THERAPY*VISIT"),
  analyse = c("BASVAL")
)

data <- data %>%
  dplyr::select(PATIENT, THERAPY, VISIT, BASVAL, THERAPY, CHANGE) %>%
  dplyr::mutate(dplyr::across(.cols = vars$id, ~ as.factor(.x))) %>%
  dplyr::arrange(dplyr::across(.cols = c(vars$id, vars$visit)))

# Use expand_locf to add rows corresponding to visits with missing outcomes to the dataset
data_full <- do.call(
  expand_locf,
  args = list(
    data = data,
    vars = c(vars$expand_vars, vars$group),
    group = vars$id,
    order = c(vars$id, vars$visit)
  ) %>%
    append(lapply(data[c(vars$id, vars$visit)], levels))
)

data_full <- data_full %>%
  dplyr::group_by(dplyr::across(vars$id)) %>%
  dplyr::mutate(!!vars$group := Filter(Negate(is.na), .data[[vars$group]])[1])

# there are duplicates - use first value
data_full <- data_full %>%
  dplyr::group_by(dplyr::across(c(vars$id, vars$group, vars$visit))) %>%
  dplyr::slice(1) %>%
  dplyr::ungroup()
# need to have a single ID column
data_full <- data_full %>%
  tidyr::unite("TMP_ID", dplyr::all_of(vars$id), sep = "_#_", remove = FALSE) %>%
  dplyr::mutate(TMP_ID = as.factor(TMP_ID))
```

#### Creating Intercurrent Event Data

Set the imputation strategy to `"MAR"` for each patient with at least one missing observation.

```{r}
#| code-fold: show

data_ice <- data_full %>%
  dplyr::arrange(dplyr::across(.cols = c("TMP_ID", vars$visit))) %>%
  dplyr::filter(is.na(.data[[missing_var]])) %>%
  dplyr::group_by(TMP_ID) %>%
  dplyr::slice(1) %>%
  dplyr::ungroup() %>%
  dplyr::select(all_of(c("TMP_ID", vars$visit))) %>%
  dplyr::mutate(strategy = "MAR")
```

## Draws

The `rbmi::draws()` function fits the imputation models and stores the corresponding parameter estimates or Bayesian posterior parameter draws.
The three main inputs to the `rbmi::draws()` function are:

-   `data` - The primary longitudinal data.frame containing the outcome variable and all covariates.
-   `data_ice` - A data.frame which specifies the first visit affected by an intercurrent event (ICE) and the imputation strategy for handling missing outcome data after the ICE. At most one ICE which is to be imputed by a non-MAR strategy is allowed per subject.
-   `method` - The statistical method used to fit the imputation models and to create imputed datasets.

#### Define Key Variables

Define the names of key variables in our dataset and the covariates included in the imputation model using `rbmi::set_vars()`.
Note that the `covariates` argument can also include interaction terms.

```{r}
#| code-fold: show

debug_mode <- FALSE

draws_vars <- rbmi::set_vars(
  outcome = missing_var,
  visit = vars$visit,
  group = vars$group,
  covariates = covariates$draws
)
draws_vars$subjid <- "TMP_ID"
```

Define which imputation method to use, then create samples for the imputation parameters by running the `draws()` function.

<!-- skip strict because of https://github.com/insightsengineering/rbmi/issues/409 -->

```{r, warning = FALSE, opts.label = "skip_test_strict"}
#| code-fold: show

draws_method <- method_bayes(seed = 123)

draws_obj <- rbmi::draws(
  data = data_full,
  data_ice = data_ice,
  vars = draws_vars,
  method = draws_method
)
```

## Impute

The next step is to use the parameters from the imputation model to generate the imputed datasets.
This is done via the `rbmi::impute()` function.
The function only has two key inputs: the imputation model output from `rbmi::draws()` and the reference groups relevant to reference-based imputation methods.
It's usage is thus:

```{r}
#| code-fold: show

impute_references <- c("DRUG" = "PLACEBO", "PLACEBO" = "PLACEBO")

impute_obj <- rbmi::impute(
  draws_obj,
  references = impute_references
)
```

## Analyze

The next step is to run the analysis model on each imputed dataset.
This is done by defining an analysis function and then calling `rbmi::analyse()` to apply this function to each imputed dataset.

<!-- skip strict because of https://github.com/insightsengineering/rbmi/issues/409 -->

```{r, opts.label = "skip_test_strict"}
#| code-fold: show
#| opts_label: "skip_test_strict" #

# Define analysis model
analyse_fun <- ancova

ref_levels <- levels(impute_obj$data$group[[1]])
names(ref_levels) <- c("ref", "alt")

analyse_obj <- rbmi::analyse(
  imputations = impute_obj,
  fun = analyse_fun,
  vars = rbmi::set_vars(
    subjid = "TMP_ID",
    outcome = missing_var,
    visit = vars$visit,
    group = vars$group,
    covariates = covariates$analyse
  )
)
```

## Pool

The `rbmi::pool()` function can be used to summarize the analysis results across multiple imputed datasets to provide an overall statistic with a standard error, confidence intervals and a p-value for the hypothesis test of the null hypothesis that the effect is equal to 0.
Using the `broom::tidy()` function the `rbmi` final results are reshaped.

<!-- skip strict because of https://github.com/insightsengineering/rbmi/issues/409 -->

```{r, opts.label = "skip_test_strict"}
#| code-fold: show

library(broom)

pool_obj <- rbmi::pool(
  results = analyse_obj,
  conf.level = 0.95,
  alternative = c("two.sided", "less", "greater"),
  type = c("percentile", "normal")
)

df <- tidy(pool_obj)
df
```

## Create Output

Finally, use functions from the `rtables` and `tern` packages to generate a nicely formatted `rtable` object.

```{r variant1, test = list(result_v1 = "result")}
result <- basic_table() %>%
  split_cols_by("group", ref_group = levels(df$group)[1]) %>%
  split_rows_by("visit", split_label = "Visit", label_pos = "topleft") %>%
  summarize_rbmi() %>%
  build_table(df)
result
```

## Data Setup

We use a publicly available example dataset from an antidepressant clinical trial of an active drug versus placebo from the `rbmi` package.
The relevant endpoint is the Hamilton 17-item depression rating scale (`HAMD17`) which was assessed at baseline and at weeks 1, 2, 4, and 6.
Study drug discontinuation occurred in 24% of subjects from the active drug and 26% of subjects from placebo.
All data after study drug discontinuation are missing and there is a single additional intermittent missing observation.

```{r setup}
#| code-fold: show
```

::::

{{< include ../../_utils/save_results.qmd >}}

{{< include ../../repro.qmd >}}
