---
title: RSPT01
subtitle: Best Overall Response
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
library(dplyr)
library(tern)

adsl <- random.cdisc.data::cadsl
adrs <- random.cdisc.data::cadrs

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adrs <- df_explicit_na(adrs)

anl_adrs <- adrs %>%
  filter(PARAMCD == "INVET") %>%
  select(STUDYID, USUBJID, PARAMCD, AVISIT, AVALC)
anl_adsl <- adsl %>%
  select(STUDYID, USUBJID, ARM, STRATA1)
```

Data pre-processing is done to label the analysis value (`AVALC`) so as to improve the readability of the output table. In addition, the *response* is binary defined (`is_rsp`, responder yes/no) depending on the analysis value. Finally, for comparison purposes, the reference arm is specified.

```{r pre-processing}
#| code-fold: show

# Pre-Processing

anl <- anl_adsl %>%
  left_join(anl_adrs, by = c("STUDYID", "USUBJID")) %>%
  mutate(rsp_lab = d_onco_rsp_label(AVALC)) %>%
  mutate(is_rsp = AVALC %in% c("CR", "PR")) %>%
  mutate(ARM = relevel(ARM, ref = "A: Drug X")) %>%
  var_relabel(ARM = "Description of Planned Arm")
```

```{r include = FALSE}
webr_code_labels <- c("setup", "pre-processing")
```

{{< include ../../_utils/webr_no_include.qmd >}}

## Output

::::::::: panel-tabset
## Standard Table

The tabulation layout is built in layers for the analysis of overall response and applied to the pre-processed dataset.

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant1, test = list(result_v1 = "result")}
lyt_01 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    table_names = "test_prop_diff"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    table_names = "est_or"
  ) %>%
  estimate_multinomial_response(var = "rsp_lab")

result <- build_table(lyt = lyt_01, df = anl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant1")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Table Selecting <br/> Sections to Display

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

Remove (or add) rows of results by removing/adding the corresponding layers from the layout. For instance, the odds-ratio row is removed by simply removing the `estimate_odds_ratio` call:

```{r variant2, test = list(result_v2 = "result")}
lyt_02 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    table_names = "test_prop"
  ) %>%
  estimate_multinomial_response(var = "rsp_lab")

result <- build_table(lyt = lyt_02, df = anl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant2")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Table Modifying Analysis Details like Type of <br/> Confidence Interval, Alpha Level, Test Statistic

The confidence level is controlled by the `conf_level` parameter to the estimation functions. Similarly, the methods for tests and confidence interval can be modified (see `?estimate_proportion_diff`).

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant3, test = list(result_v3 = "result")}
conf_level <- 0.90
lyt_03 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    conf_level = conf_level,
    method = "clopper-pearson",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    conf_level = conf_level,
    method = "ha",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    method = "fisher",
    table_names = "test_prop"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    conf_level = conf_level,
    table_names = "est_or"
  ) %>%
  estimate_multinomial_response(
    var = "rsp_lab",
    conf_level = conf_level,
    method = "clopper-pearson"
  )

result <- build_table(lyt = lyt_03, df = anl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant3")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Table with <br/> Stratified Analysis

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

The stratified analysis section can be added by defining the analyses needed with `control_binary_comparison` for the argument `strat_analysis` and identifying the stratification variables to use.

```{r variant4, test = list(result_v4 = "result")}
strata <- "STRATA1"
lyt_04 <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM", ref_group = "A: Drug X") %>%
  estimate_proportion(
    vars = "is_rsp",
    table_names = "est_prop"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Unstratified Analysis",
    table_names = "est_prop_diff"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    table_names = "test_prop"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    table_names = "est_or"
  ) %>%
  estimate_proportion_diff(
    vars = "is_rsp",
    show_labels = "visible",
    var_labels = "Stratified Analysis",
    method = "cmh",
    variables = list(strata = strata),
    table_names = "est_prop_diff_strat"
  ) %>%
  test_proportion_diff(
    vars = "is_rsp",
    method = "cmh",
    variables = list(strata = strata),
    table_names = "test_prop_strat"
  ) %>%
  estimate_odds_ratio(
    vars = "is_rsp",
    variables = list(strata = strata, arm = "ARM"),
    table_names = "est_or_strat"
  ) %>%
  estimate_multinomial_response(var = "rsp_lab")

result <- build_table(lyt = lyt_04, df = anl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant4")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Table Modifying the Definition <br/> of Overall Response

The definition of *responders* is realized during the pre-processing step. The layout does not need to be modified and can be reused.

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant5, test = list(result_v5 = "result")}
anl <- anl_adsl %>%
  left_join(anl_adrs, by = c("STUDYID", "USUBJID")) %>%
  mutate(rsp_lab = d_onco_rsp_label(AVALC)) %>%
  mutate(is_rsp = AVALC == "CR") %>%
  mutate(ARM = relevel(ARM, ref = "A: Drug X")) %>%
  var_relabel(ARM = "Description of Planned Arm")

result <- build_table(lyt = lyt_01, df = anl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant5")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Table Defining New <br/> Sections to Display

Similarly to in the previous tab, redefinition or relabeling of the result is a pre-processing step and the original table layout can be reused.

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant6, test = list(result_v6 = "result")}
anl <- anl_adsl %>%
  left_join(anl_adrs, by = c("STUDYID", "USUBJID")) %>%
  mutate(rsp_lab = as.character(d_onco_rsp_label(AVALC))) %>%
  mutate(
    rsp_lab = case_when(
      rsp_lab == "Complete Response (CR)" ~ "No Progression",
      rsp_lab == "Partial Response (PR)" ~ "No Progression",
      rsp_lab == "Stable Disease (SD)" ~ "No Progression",
      TRUE ~ rsp_lab
    )
  ) %>%
  mutate(is_rsp = rsp_lab %in% "No Progression") %>%
  mutate(ARM = relevel(ARM, ref = "A: Drug X")) %>%
  var_relabel(ARM = "Description of Planned Arm")

result <- build_table(lyt = lyt_01, df = anl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant6")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Data Setup

```{r setup}
#| code-fold: show
```
:::::::::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

```{r teal, opts.label = c("skip_if_testing", "app")}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(dplyr)

  ADSL <- random.cdisc.data::cadsl
  ADRS <- random.cdisc.data::cadrs

  # Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
  ADSL <- df_explicit_na(ADSL)
  ADRS <- df_explicit_na(ADRS)

  ADSL <- ADSL %>%
    mutate(Dum_ARM = factor(rep("Single ARM", nrow(.))))
  ADRS <- ADRS %>%
    mutate(Dum_ARM = factor(rep("Single ARM", nrow(.))))
})
datanames <- c("ADSL", "ADRS")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADRS <- data[["ADRS"]]
arm_ref_comp <- list(
  ACTARMCD = list(
    ref = "ARM B",
    comp = c("ARM A", "ARM C")
  ),
  ARM = list(
    ref = "B: Placebo",
    comp = c("A: Drug X", "C: Combination")
  )
)

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_t_binary_outcome(
      label = "Responders",
      dataname = "ADRS",
      paramcd = choices_selected(
        choices = value_choices(ADRS, "PARAMCD", "PARAM"),
        selected = "BESRSPI"
      ),
      arm_var = choices_selected(
        choices = variable_choices(ADRS, c("ARM", "ARMCD", "ACTARMCD", "Dum_ARM")),
        selected = "ARM"
      ),
      arm_ref_comp = arm_ref_comp,
      strata_var = choices_selected(
        choices = variable_choices(ADRS, c("SEX", "BMRKR2")),
        select = NULL
      ),
      rsp_table = TRUE
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../_utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}