---
title: MMRMG01
subtitle: Plots for Mixed-Effect Model Repeated Measures Analysis
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
library(dplyr)
library(tern.mmrm)
library(nestcolor)

adsl <- random.cdisc.data::cadsl
adqs <- random.cdisc.data::cadqs

adqs_f <- adqs %>%
  dplyr::filter(PARAMCD == "FKSI-FWB" & !AVISIT %in% c("BASELINE")) %>%
  droplevels() %>%
  dplyr::mutate(ARM = factor(ARM, levels = c("B: Placebo", "A: Drug X", "C: Combination"))) %>%
  dplyr::mutate(AVISITN = rank(AVISITN) %>% as.factor() %>% as.numeric() %>% as.factor())

mmrm_results <- fit_mmrm(
  vars = list(
    response = "AVAL",
    covariates = c("STRATA2"),
    id = "USUBJID",
    arm = "ARM",
    visit = "AVISIT"
  ),
  data = adqs_f
)
```

Given an MMRM fitted with `s_mmrm`, `g_mmrm_lsmeans` displays for each visit the adjusted means within group and/or difference in adjusted means between groups. `g_mmrm_diagnostic` displays marginal residual plots for evaluating model fit.

## Least Squares Means: <br/> Estimates Within Groups

### Considering the treatment variable in the model

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->

```{r plot1, dev.args = list(pointsize = 6), test = list(plot_v1 = "plot"), opts.label = "skip_test_strict"}
plot <- g_mmrm_lsmeans(
  mmrm_results,
  select = "estimates",
  xlab = "Visit"
)
plot
```

### Considering the treatment variable in the model, with lines

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->

```{r plot2, dev.args = list(pointsize = 6), test = list(plot_v2 = "plot"), opts.label = "skip_test_strict"}
plot <- g_mmrm_lsmeans(
  mmrm_results,
  select = "estimates",
  xlab = "Visit",
  show_lines = TRUE
)
plot
```

### Considering the treatment variable in the model, with statistics table

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->

```{r plot3, dev.args = list(pointsize = 6), fig.height = 7, test = list(plot_v3 = "plot"), opts.label = "skip_test_strict"}
plot <- g_mmrm_lsmeans(
  mmrm_results,
  select = "estimates",
  xlab = "Visit",
  table_stats = c("n", "estimate", "se", "ci"),
  table_font_size = 4,
  table_rel_height = 0.6
)
plot
```

## Least Squares Means: <br/> Contrasts Between Groups

Users can choose to display both estimates and contrasts together by running `g_mmrm_lsmeans(mmrm_results)`.

<!-- skip strict because of https://github.com/r-lib/gtable/pull/94 -->

```{r plot4, dev.args = list(pointsize = 6), test = list(plot_v4 = "plot"), opts.label = "skip_test_strict"}
plot <- g_mmrm_lsmeans(
  mmrm_results,
  select = "contrasts",
  titles = c(contrasts = "Contrasts of FKSI-FWB means"),
  xlab = "Visit"
)
plot
```

## Model Diagnostics: Marginal <br/> Fitted Values vs. Residuals

```{r plot5, dev.args = list(pointsize = 6), test = list(plot_v5 = "plot")}
plot <- g_mmrm_diagnostic(mmrm_results)
plot
```

## Model Diagnostics: Normality <br/> of Marginal Residuals

```{r plot6, dev.args = list(pointsize = 6), test = list(plot_v6 = "plot")}
plot <- g_mmrm_diagnostic(mmrm_results, type = "q-q-residual")
plot
```

## Data Setup and <br/> Model Fitting

```{r setup}
#| code-fold: show
```

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

```{r teal, opts.label = c("skip_if_testing", "app")}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  library(dplyr)

  ADSL <- random.cdisc.data::cadsl
  ADQS <- random.cdisc.data::cadqs %>%
    filter(ABLFL != "Y" & ABLFL2 != "Y") %>%
    filter(AVISIT %in% c("WEEK 1 DAY 8", "WEEK 2 DAY 15", "WEEK 3 DAY 22")) %>%
    mutate(
      AVISIT = as.factor(AVISIT),
      AVISITN = rank(AVISITN) %>%
        as.factor() %>%
        as.numeric() %>%
        as.factor() # making consecutive numeric factor
    )
})
datanames <- c("ADSL", "ADQS")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

arm_ref_comp <- list(
  ARMCD = list(
    ref = "ARM A",
    comp = c("ARM B", "ARM C")
  )
)

## Reusable Configuration For Modules
ADQS <- data[["ADQS"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_a_mmrm(
      label = "MMRM",
      dataname = "ADQS",
      aval_var = choices_selected(c("AVAL", "CHG"), "AVAL"),
      id_var = choices_selected(c("USUBJID", "SUBJID"), "USUBJID"),
      arm_var = choices_selected(c("ARM", "ARMCD"), "ARMCD"),
      visit_var = choices_selected(c("AVISIT", "AVISITN"), "AVISIT"),
      arm_ref_comp = arm_ref_comp,
      paramcd = choices_selected(
        choices = value_choices(ADQS, "PARAMCD", "PARAM"),
        selected = "FKSI-FWB"
      ),
      cov_var = choices_selected(c("BASE", "AGE", "SEX", "BASE:AVISIT"), NULL),
      conf_level = choices_selected(c(0.95, 0.9, 0.8), 0.95)
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../repro.qmd >}}