---
title: AET06
subtitle: Adverse Events by Baseline Characteristic
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

:::: {.panel-tabset}
## Data Setup

```{r setup, message = FALSE}
#| code-fold: show

library(dplyr)
library(tern)

adsl <- random.cdisc.data::cadsl
adae <- random.cdisc.data::cadae
adsub <- random.cdisc.data::cadsub

adsub_bmi <- adsub %>%
  filter(PARAMCD == "BBMISI") %>%
  select(STUDYID, USUBJID, AVALCAT1) %>%
  mutate(
    AVALCAT1 = factor(AVALCAT1, levels = c("<18.5", "18.5 - 24.9", "25 - 29.9", ">30"))
  )

adsl <- adsl %>%
  mutate(
    RACE1 = case_when(
      RACE == "WHITE" ~ "WHITE",
      TRUE ~ "NON-WHITE"
    ),
    RACE1 = factor(
      RACE1,
      levels = c("WHITE", "NON-WHITE")
    )
  ) %>%
  left_join(
    y = adsub_bmi,
    by = c("STUDYID", "USUBJID")
  )

adae_labels <- var_labels(adae)

adae <- adae %>%
  mutate(
    RACE1 = case_when(
      RACE == "WHITE" ~ "WHITE",
      TRUE ~ "NON-WHITE"
    ),
    RACE1 = factor(
      RACE1,
      levels = c("WHITE", "NON-WHITE")
    )
  ) %>%
  left_join(
    y = adsub_bmi,
    by = c("STUDYID", "USUBJID")
  )

var_labels(adae) <- c(adae_labels, "RACE1" = "RACE WHITE/NON-WHITE", "AVALCAT1" = "Baseline BMI Category")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adae <- df_explicit_na(adae)
adsub_bmi <- df_explicit_na(adsub_bmi)
```

## Adverse Events by Sex

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant1, test = list(result_v1 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  split_cols_by("SEX") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = -1L) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt = lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  # Sorted by decreasing frequency across all groups by System Organ Class and Preferred Term.
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

result
```

`r webr_code_labels <- c("setup", "variant1")` {{< include ../../_utils/webr.qmd >}}
:::

## Adverse Events by Other Baseline <br/> Characteristics (e.g. Biomarker Group)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant2, test = list(result_v2 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  split_cols_by("BMRKR2") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = -1L) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt = lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  # Sorted by decreasing frequency across all groups by System Organ Class and Preferred Term.
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

result
```

`r webr_code_labels <- c("setup", "variant2")` {{< include ../../_utils/webr.qmd >}}
:::

## Adverse Events by Baseline Characteristic <br/> from ADSUB (e.g. BMI Category)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant3, test = list(result_v3 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  split_cols_by("AVALCAT1") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = -1L) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt = lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  # Sorted by decreasing frequency across all groups by System Organ Class and Preferred Term.
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

result
```

`r webr_code_labels <- c("setup", "variant3")` {{< include ../../_utils/webr.qmd >}}
:::

## Adverse Events by Regrouped Baseline <br/> Characteristics (e.g. Race)

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant4, test = list(result_v4 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  split_cols_by("RACE1") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences(vars = "AEDECOD", .indent_mods = -1L) %>%
  append_varlabels(adae, "AEDECOD", indent = 1L)

result <- build_table(
  lyt = lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  prune_table() %>%
  # Sorted by decreasing frequency across all groups by System Organ Class and Preferred Term.
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

result
```

`r webr_code_labels <- c("setup", "variant4")` {{< include ../../_utils/webr.qmd >}}
:::

## Adverse Events by Baseline Characteristics <br/> (e.g. Sex) Including High-Level Terms

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant5, test = list(result_v5 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  split_cols_by("SEX") %>%
  analyze_num_patients(
    vars = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Overall total number of events"
    )
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = FALSE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  split_rows_by(
    "AEHLT",
    child_labels = "visible",
    indent_mod = -1L,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEHLT)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = c("unique", "nonunique"),
    .labels = c(
      unique = "Total number of patients with at least one adverse event",
      nonunique = "Total number of events"
    )
  ) %>%
  count_occurrences("AEDECOD", .indent_mods = -1L) %>%
  append_varlabels(adae, "AEDECOD", indent = 2L)

result <- build_table(
  lyt = lyt,
  df = adae,
  alt_counts_df = adsl
) %>%
  prune_table() %>%
  # Sorted by decreasing frequency across all groups by System Organ Class, High-Level Term and Preferred Term.
  sort_at_path(
    path = c("AEBODSYS"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEHLT"),
    scorefun = cont_n_allcols
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEHLT", "*", "AEDECOD"),
    scorefun = score_occurrences
  )

result
```

`r webr_code_labels <- c("setup", "variant5")` {{< include ../../_utils/webr.qmd >}}
:::

{{< include ../../_utils/save_results.qmd >}}

## `teal` App

::: {.panel-tabset .nav-justified}
## {{< fa regular file-lines fa-sm fa-fw >}} Preview

```{r teal, opts.label = c("skip_if_testing", "app")}
library(teal.modules.clinical)

## Data reproducible code
data <- teal_data()
data <- within(data, {
  ADSL <- random.cdisc.data::cadsl
  ADAE <- random.cdisc.data::cadae

  # Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
  ADSL <- df_explicit_na(ADSL)
  ADAE <- df_explicit_na(ADAE)
})
datanames <- c("ADSL", "ADAE")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]

## Reusable Configuration For Modules
ADAE <- data[["ADAE"]]

## Setup App
app <- init(
  data = data,
  modules = modules(
    tm_t_events(
      label = "Adverse Event Table",
      dataname = "ADAE",
      arm_var = choices_selected(c("ARM", "ARMCD", "SEX"), c("ARM", "SEX")),
      llt = choices_selected(
        choices = variable_choices(ADAE, c("AETERM", "AEDECOD")),
        selected = c("AEDECOD")
      ),
      hlt = choices_selected(
        choices = variable_choices(ADAE, c("AEBODSYS", "AESOC")),
        selected = "AEBODSYS"
      ),
      add_total = TRUE,
      event_type = "adverse event"
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../_utils/shinylive.qmd >}}
:::

{{< include ../../repro.qmd >}}
::::
