---
title: AET04
subtitle: Adverse Events by Highest NCI CTCAE Grade
---

------------------------------------------------------------------------

{{< include ../../test_utils.qmd >}}

::: panel-tabset
## Data Setup

```{r setup, message=FALSE}
#| code-fold: show

library(dplyr)
library(tern)
library(scda)

adsl <- synthetic_cdisc_dataset("latest", "adsl")
adae <- synthetic_cdisc_dataset("latest", "adae")

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl) %>% dplyr::filter(TRT01A != "<Missing>")
adae <- df_explicit_na(adae) %>% dplyr::filter(ANL01FL == "Y")

# Pre-Processing

gr_grp <- list(
  "- Any Grade -" = c("1", "2", "3", "4", "5"),
  "Grade 1-2" = c("1", "2"),
  "Grade 3-4" = c("3", "4"),
  "Grade 5" = "5"
)

# Raw table used by variant 8/10
raw_table <- basic_table() %>%
  split_cols_by("ACTARM") %>%
  add_colcounts() %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by("AEBODSYS",
    split_fun = trim_levels_in_group("AETOXGR"),
    child_labels = "visible", nested = TRUE
  ) %>%
  summarize_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by("AEDECOD",
    split_fun = trim_levels_in_group("AETOXGR"),
    child_labels = "visible", nested = TRUE
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "- Any Grade -"
  ) %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp[-1],
    .indent_mods = -1L
  ) %>%
  build_table(adae, alt_counts_df = adsl) %>%
  sort_at_path(
    path = "AEBODSYS",
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  )

# Helper function to avoid filtering also the first part of the table, where general information is given.
my_row_condition <- function(row_fnc_condition) {
  function(table_row) {
    if (indent_mod(table_row) == 0) {
      return(TRUE)
    } else {
      row_fnc_condition(table_row)
    }
  }
}
```

## Standard Table

```{r variant1, test = list(result_v1 = "result")}
split_fun <- trim_levels_in_group

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEDECOD",
    child_labels = "visible",
    nested = TRUE,
    indent_mod = -1L,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae$AEDECOD)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "- Any Grade -"
  ) %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp[-1],
    .indent_mods = -1L
  ) %>%
  append_varlabels(adae, "AETOXGR", indent = 2L)

result <- lyt %>%
  build_table(adae, alt_counts_df = adsl) %>%
  sort_at_path(
    path = "AEBODSYS",
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  )

result
```

## Table with Fill-In <br/> of Treatment Groups

```{r variant2, test = list(result_v2 = "result")}
adae2 <- adae %>%
  dplyr::filter(ACTARM == "A: Drug X")

split_fun <- trim_levels_in_group

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae2$AEBODSYS)
  ) %>%
  summarize_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEDECOD",
    child_labels = "visible",
    nested = TRUE,
    indent_mod = -1L,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae2$AEDECOD)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "- Any Grade -"
  ) %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp[-1],
    .indent_mods = -1L
  ) %>%
  append_varlabels(adae, "AETOXGR", indent = 2L)

result <- lyt %>%
  build_table(adae2, alt_counts_df = adsl) %>%
  sort_at_path(
    path = "AEBODSYS",
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  )

result
```

## Table with Fill-In <br/> of Grades

```{r variant3, test = list(result_v3 = "result")}
split_fun <- drop_split_levels

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEDECOD",
    child_labels = "visible",
    nested = TRUE,
    indent_mod = -1L,
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae$AEDECOD)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "- Any Grade -"
  ) %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp[-1],
    .indent_mods = -1L
  ) %>%
  append_varlabels(adae, "AETOXGR", indent = 2L)

result <- lyt %>%
  build_table(adae, alt_counts_df = adsl) %>%
  sort_at_path(
    path = "AEBODSYS",
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  )

result
```

## Table with Collapsing <br/> of Grades

```{r variant4, test = list(result_v4 = "result")}
gr_grp1 <- list(
  "- Any Grade -" = c("1", "2", "3", "4", "5"),
  "Grade 1-2" = c("1", "2"),
  "Grade 3-5" = c("3", "4", "5")
)

split_fun <- trim_levels_in_group

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp1
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae$AEBODSYS)
  ) %>%
  summarize_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp1
  ) %>%
  split_rows_by(
    "AEDECOD",
    child_labels = "visible",
    nested = TRUE,
    indent_mod = -1L,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae$AEDECOD)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "- Any Grade -"
  ) %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp1[-1],
    .indent_mods = -1L
  ) %>%
  append_varlabels(adae, "AETOXGR", indent = 2L)

result <- lyt %>%
  build_table(adae, alt_counts_df = adsl) %>%
  prune_table() %>%
  sort_at_path(
    path = "AEBODSYS",
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  )

result
```

## Table Using Worst Grade <br/> Flags from ADAE

```{r variant5}
#| code-fold: show

# No Worst Grade Flags found in the ADAE data set.
```

## Table with an Incidence Rate <br/> $\geq$ 40%, Totals Restricted

```{r variant6, test = list(result_v6 = "result")}
# Simple wrapper to return subset ADAE to a threshold of xx%.
get_adae_trimmed <- function(adsl, adae, cutoff_rate) {
  n_per_arm <- adsl %>%
    dplyr::count(ACTARM)

  anl_terms <- adae %>%
    dplyr::group_by(ACTARM, AEBODSYS, AEDECOD) %>%
    dplyr::count(
      unique_terms = n_distinct(USUBJID)
    ) %>%
    dplyr::select(-n) %>%
    dplyr::ungroup()

  anl_terms <- dplyr::left_join(
    anl_terms,
    n_per_arm,
    by = "ACTARM"
  ) %>%
    dplyr::mutate(
      ae_rate = unique_terms / n
    ) %>%
    dplyr::filter(ae_rate >= cutoff_rate) %>%
    dplyr::select(AEDECOD) %>%
    unique()

  anl <- dplyr::left_join(
    anl_terms,
    adae,
    by = "AEDECOD"
  )
  anl
}

adae6 <- get_adae_trimmed(adsl, adae, cutoff_rate = 0.4)

split_fun <- trim_levels_in_group

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ACTARM") %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEBODSYS",
    child_labels = "visible",
    nested = TRUE,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae6$AEBODSYS)
  ) %>%
  summarize_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp
  ) %>%
  split_rows_by(
    "AEDECOD",
    child_labels = "visible",
    nested = TRUE,
    indent_mod = -1L,
    split_fun = split_fun("AETOXGR"),
    label_pos = "topleft",
    split_label = obj_label(adae6$AEDECOD)
  ) %>%
  summarize_num_patients(
    var = "USUBJID",
    .stats = "unique",
    .labels = "- Any Grade -"
  ) %>%
  count_occurrences_by_grade(
    var = "AETOXGR",
    grade_groups = gr_grp[-1],
    .indent_mods = -1L
  ) %>%
  append_varlabels(adae, "AETOXGR", indent = 2L)

result <- lyt %>%
  build_table(adae6, alt_counts_df = adsl) %>%
  prune_table() %>%
  sort_at_path(
    path = "AEBODSYS",
    scorefun = cont_n_allcols,
    decreasing = TRUE
  ) %>%
  sort_at_path(
    path = c("AEBODSYS", "*", "AEDECOD"),
    scorefun = cont_n_allcols,
    decreasing = TRUE
  )

result
```

## Table with an Incidence Rate <br/> $\geq$ X%, Totals Unrestricted

```{r variant7}
#| code-fold: show

# Variant 7 was not created.
# With this variant, the SOC level is not trimmed (even if there are no terms left).
```

## Table with an Incidence <br/> Rate $\geq$ 58 Patients

```{r variant8, test = list(result_v8 = "result")}
cutoff <- 58L
row_condition <- has_count_in_any_col(atleast = cutoff, col_names = levels(adsl$ACTARM))

result <- prune_table(raw_table, keep_content_rows(my_row_condition(row_condition)))

result
```

## Table with a Difference in <br/> Incidence Rate $\geq$ 10%

```{r variant9, test = list(result_v9 = "result")}
cutoff <- 0.1
row_condition <- has_fractions_difference(atleast = cutoff, col_names = levels(adsl$ACTARM))

result <- prune_table(raw_table, keep_content_rows(my_row_condition(row_condition)))

result
```

## Table with an Incidence Rate <br/> $\geq$ 5%, SOCs \< 5% Removed

```{r variant10}
#| code-fold: show

# Variant 10 was not done
# With this variant, SOC levels above the threshold are still in the table even if
# there are no terms left.
```

## Table with an Incidence Rate $\geq$ 40%, <br/> All SOCs w/o Preferred Terms Removed

```{r variant11, test = list(result_v11 = "result")}
cutoff <- 0.4
row_condition <- has_fraction_in_any_col(atleast = cutoff, col_names = levels(adsl$ACTARM))

result <- prune_table(raw_table, keep_content_rows(my_row_condition(row_condition)))

result
```

```{r save testenv, include=FALSE, opts.label='include_if_testing'}
#| renv.ignore: TRUE
dirname <- Sys.getenv("QUARTO_TESTTHAT_DATA_PATH")
filename <- paste0(gsub("\\.rmarkdown$", "", knitr::current_input()), ".rds")
filepath <- file.path(dirname, filename)
file.create(filepath)
saveRDS(tenv, filepath)
```

## Teal

```{r teal, message=FALSE, opts.label='skip_if_testing'}
#| screenshot.opts = list(delay = 12)

library(dplyr)
library(scda)
library(teal.modules.clinical)
library(tern)

adsl <- synthetic_cdisc_dataset("latest", "adsl")
adae <- synthetic_cdisc_dataset("latest", "adae")

app <- init(
  data = cdisc_data(
    cdisc_dataset("ADSL", adsl),
    cdisc_dataset("ADAE", adae),
    code = 'ADSL <- synthetic_cdisc_dataset("latest", "adsl")
    ADAE <- synthetic_cdisc_dataset("latest", "adae")',
    check = TRUE
  ),
  modules = modules(
    tm_t_events_by_grade(
      label = "Adverse Events by Grade Table",
      dataname = "ADAE",
      arm_var = choices_selected(c("ARM", "ARMCD"), "ARM"),
      llt = choices_selected(
        choices = variable_choices(adae, c("AETERM", "AEDECOD")),
        selected = c("AEDECOD")
      ),
      hlt = choices_selected(
        choices = variable_choices(adae, c("AEBODSYS", "AESOC")),
        selected = "AEBODSYS"
      ),
      grade = choices_selected(
        choices = variable_choices(adae, c("AETOXGR", "AESEV")),
        selected = "AETOXGR"
      )
    )
  )
)

shinyApp(app$ui, app$server)
```

{{< include ../../si.qmd >}}
:::
