---
title: DISCLOSUREST01
subtitle: Disclosures Outputs
---

------------------------------------------------------------------------

{{< include ../../_utils/envir_hook.qmd >}}

::::::::: panel-tabset
## Patient Disposition Table <br/> (CTgov & EudraCT)

#### Data Setup

```{r setup1, message = FALSE}
library(tern)
library(dplyr)

adsl <- random.cdisc.data::cadsl

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)

set.seed(1, kind = "Mersenne-Twister")

# Add additional disposition variables to adsl.
adsl0 <- adsl %>%
  mutate(
    STSTFL = case_when(
      is.na(RANDDT) ~ "N",
      TRUE ~ "Y"
    ) %>% as.factor(),
    COMPSTUD = sample(
      c("Y", "N"),
      size = nrow(adsl),
      replace = TRUE
    ) %>% as.factor(),
    STUDONS = sample(
      c("Alive: On Treatment", "Alive: In Follow-up", NA),
      size = nrow(adsl),
      replace = TRUE
    ) %>% as.factor(),
    STDDRS = sample(
      c(
        "Death", "Lost To Follow-Up",
        "Protocol Violation", "Withdrawal By Subject",
        "Other"
      ),
      size = nrow(adsl),
      replace = TRUE
    ) %>% as.factor(),
    GOTTRT = ifelse(!is.na(ACTARMCD), "Y", "N") %>%
      as.factor(),
    DISTRTFL = sample(
      c("Y", "N"),
      size = nrow(adsl),
      replace = TRUE
    ) %>% as.factor(),
    TRTDRS = sample(
      c(
        "ADVERSE EVENT", "PROGRESSIVE DISEASE",
        "PHYSICIAN DECISION", "LACK OF EFFICACY",
        "OTHER"
      ),
      size = nrow(adsl),
      replace = TRUE
    ) %>% as.factor(),
    STUDONS = case_when(COMPSTUD == "N" ~ STUDONS),
    STDDRS = case_when(COMPSTUD == "N" & is.na(STUDONS) ~ STDDRS),
    DISSTDFL = case_when(!is.na(STDDRS) ~ "Y"),
    DISTRTFL = case_when(GOTTRT == "Y" ~ DISTRTFL),
    TRTDRS = case_when(DISTRTFL == "Y" ~ TRTDRS),
    DRSCAT = case_when(
      TRTDRS %in% c("ADVERSE EVENT", "PHYSICIAN DECISION") ~ "Safety",
      !is.na(TRTDRS) ~ "Other"
    )
  ) %>%
  var_relabel(
    STSTFL = "Started Study",
    COMPSTUD = "Complete Study",
    STUDONS = "On-study Status",
    DISSTDFL = "Discontinued Study",
    STDDRS = "Reason for Study Discontinuation",
    GOTTRT = "Received Treatment",
    DISTRTFL = "Discontinued Treatment",
    TRTDRS = "Reason for Treatment Discontinuation",
    DRSCAT = "Subcategory for Treatment Discontinuation"
  )
```

```{r include = FALSE}
webr_code_labels <- c("setup1")
```

{{< include ../../_utils/webr_no_include.qmd >}}

#### Patient Disposition Table

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant1, test = list(result_v1 = "result")}
# Define the split function
split_fun <- keep_split_levels("Y")

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(
    "ARM",
    split_fun = add_overall_level("All Patients", first = FALSE)
  ) %>%
  count_values("STSTFL",
    values = "Y",
    .labels = c(count_fraction = "Started Study")
  ) %>%
  count_values("COMPSTUD",
    values = "Y",
    .labels = c(count_fraction = "Completed Study")
  ) %>%
  split_rows_by(
    "DISSTDFL",
    split_fun = split_fun
  ) %>%
  summarize_row_groups(label_fstr = "Discontinued Study") %>%
  analyze_vars(
    "STDDRS",
    .stats = "count_fraction"
  ) %>%
  append_topleft("Status")

result <- build_table(lyt = lyt, df = adsl0)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant1")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Demographic Table <br/> (CTgov & EudraCT)

#### Data Setup

```{r setup2, message = FALSE}
library(tern)
library(dplyr)

adsl <- random.cdisc.data::cadsl

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)

# Change description in variable SEX.
adsl <- adsl %>%
  mutate(
    SEX = factor(case_when(
      SEX == "M" ~ "Male",
      SEX == "F" ~ "Female",
      SEX == "U" ~ "Unknown",
      SEX == "UNDIFFERENTIATED" ~ "Undifferentiated"
    )),
    AGEGRP = factor(
      case_when(
        between(AGE, 18, 40) ~ "18-40",
        between(AGE, 41, 64) ~ "41-64",
        AGE > 64 ~ ">=65"
      ),
      levels = c("18-40", "41-64", ">=65")
    )
  )
```

```{r include = FALSE}
webr_code_labels <- c("setup2")
```

{{< include ../../_utils/webr_no_include.qmd >}}

#### Demographic Table

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant2, test = list(result_v2 = "result")}
vars <- c("AGE", "AGEGRP", "SEX", "RACE", "ETHNIC")
var_labels <- c("Age (yr)", "Age group", "Sex", "Race", "Ethnicity")

lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by(var = "ARM") %>%
  add_overall_col("All Patients") %>%
  analyze_vars(
    vars = vars,
    var_labels = var_labels
  )

result <- build_table(lyt = lyt, df = adsl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant2")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Enrollment by Country <br/> Table (EudraCT only)

#### Data Setup

```{r setup3, message = FALSE}
library(tern)
library(dplyr)

adsl <- random.cdisc.data::cadsl
adsl_labels <- var_labels(adsl)

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)

adsl <- adsl %>%
  mutate(COUNTRY = droplevels(COUNTRY)) %>%
  arrange(REGION1, COUNTRY)

var_labels(adsl) <- c(adsl_labels)
```

```{r include = FALSE}
webr_code_labels <- c("setup3")
```

{{< include ../../_utils/webr_no_include.qmd >}}

#### Enrollment by Country Table

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant3, test = list(result_v3 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col(label = "All Patients") %>%
  analyze_vars("COUNTRY") %>%
  append_varlabels(adsl, "COUNTRY")

result <- build_table(lyt, adsl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant3")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Non-Serious Adverse Events <br/> Reported in $\geq$ 5% of Patients in Any <br/> Treatment Group (CTgov & EudraCT)

#### Data Setup

Trimming function `get_adae_trimmed` is defined to filter AEs with greater than 5% incidence rate.

```{r setup4, message = FALSE}
library(tern)
library(dplyr)

adsl <- random.cdisc.data::cadsl
adae <- random.cdisc.data::cadae

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adsl <- df_explicit_na(adsl)
adae <- df_explicit_na(adae)

adae_nonser <- adae %>% filter(AESER != "Y", SAFFL == "Y")

get_adae_trimmed <- function(adsl, adae, cutoff_rate) {
  n_per_arm <- adsl %>%
    dplyr::count(ARM)

  anl_terms <- adae %>%
    dplyr::group_by(ARM, AEBODSYS, AEDECOD) %>%
    dplyr::count(
      unique_terms = n_distinct(USUBJID)
    ) %>%
    dplyr::select(-n) %>%
    dplyr::ungroup()

  anl_terms <- dplyr::left_join(
    anl_terms,
    n_per_arm,
    by = "ARM"
  ) %>%
    dplyr::mutate(
      ae_rate = unique_terms / n
    ) %>%
    dplyr::filter(ae_rate >= cutoff_rate) %>%
    dplyr::select(AEDECOD) %>%
    unique()

  anl <- dplyr::left_join(
    anl_terms,
    adae,
    by = "AEDECOD"
  )
  anl
}

adae_trim <- get_adae_trimmed(adsl, adae_nonser, cutoff_rate = 0.05)
```

```{r include = FALSE}
webr_code_labels <- c("setup4")
```

{{< include ../../_utils/webr_no_include.qmd >}}

#### Non-Serious Adverse Events Report in $\geq$ 5% of Patients in Any Treatment Group

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant4, test = list(result_v4 = "result")}
# Define the split function
split_fun <- drop_split_levels

lyt <- basic_table() %>%
  split_cols_by("ARM") %>%
  summarize_patients_events_in_cols(
    custom_label = "Total number of patients with at least one non-serious adverse event occuring at a relative frequency of >=5%" # nolint: line_length.
  ) %>%
  split_rows_by("AEBODSYS",
    nested = FALSE,
    split_fun = split_fun,
    indent_mod = -1L,
    label_pos = "topleft",
    split_label = obj_label(adae_trim$AEBODSYS)
  ) %>%
  split_rows_by("AEDECOD", split_fun = split_fun, label_pos = "topleft", split_label = obj_label(adae_trim$AEDECOD)) %>%
  summarize_patients_events_in_cols(
    col_split = FALSE
  )

result <- build_table(lyt, adae_trim)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant4")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Serious Adverse Events, Fatal SAEs <br/> and SAEs Related to Study Medication, <br/> by Treatment Group (CTgov & EudraCT)

For illustrative purposes, the `adae` data is filtered by arm "A: Drug X" here.

#### Data Setup

```{r setup5, message = FALSE}
library(tern)
library(dplyr)

adae <- random.cdisc.data::cadae

# Ensure character variables are converted to factors and empty strings and NAs are explicit missing levels.
adae <- df_explicit_na(adae)

adae_serious <- adae %>% filter(AESER == "Y", SAFFL == "Y")
adae_serious_arm <- adae_serious %>% filter(ARM == "A: Drug X")

filters_list <- list(
  related = with_label(c(AEREL = "Y"), "Events (Related)"),
  fatal = with_label(c(AESDTH = "Y"), "Events (Fatal)"),
  fatal_related = with_label(c(AEREL = "Y", AESDTH = "Y"), "Events (Fatal & Related)")
)
```

```{r include = FALSE}
webr_code_labels <- c("setup5")
```

{{< include ../../_utils/webr_no_include.qmd >}}

#### Serious Adverse Events, Fatal SAEs and SAEs Related to Study Medication, by Treatment Group

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant5, test = list(result_v5 = "result")}
# Define the split function
split_fun <- drop_split_levels

lyt <- basic_table() %>%
  summarize_patients_events_in_cols(
    filters_list = filters_list,
    empty_stats = c("all", "related", "fatal", "fatal_related"),
    custom_label = "Total number of patients with at least one serious adverse event"
  ) %>%
  split_rows_by("AEBODSYS",
    nested = FALSE,
    split_fun = split_fun,
    indent_mod = -1L,
    label_pos = "topleft",
    split_label = obj_label(adae_trim$AEBODSYS)
  ) %>%
  split_rows_by("AEDECOD",
    split_fun = split_fun,
    label_pos = "topleft",
    split_label = obj_label(adae_serious_arm$AEDECOD)
  ) %>%
  summarize_patients_events_in_cols(
    filters_list = filters_list,
    col_split = FALSE
  )

result <- build_table(lyt, adae_serious_arm)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant5")
```

{{< include ../../_utils/webr.qmd >}}
:::

## Death Table <br/> (EudraCT only)

#### Data Setup

```{r setup6, message = FALSE}
library(tern)
library(dplyr)

adsl <- random.cdisc.data::cadsl
adae <- random.cdisc.data::cadae
```

```{r include = FALSE}
webr_code_labels <- c("setup6")
```

{{< include ../../_utils/webr_no_include.qmd >}}

#### Death Table

::: {.panel-tabset .nav-justified group="webr"}
## {{< fa regular file-lines sm fw >}} Preview

```{r variant6, test = list(result_v6 = "result")}
lyt <- basic_table(show_colcounts = TRUE) %>%
  split_cols_by("ARM") %>%
  add_overall_col("All Patients") %>%
  count_patients_with_event(
    "USUBJID",
    filters = c("AESDTH" = "Y"),
    .labels = c(count_fraction = "Total Number of Deaths"),
    .formats = c(count_fraction = "xx (xx.xx%)")
  )
result <- build_table(lyt, adae, alt_counts_df = adsl)
result
```

```{r include = FALSE}
webr_code_labels <- c("variant6")
```

{{< include ../../_utils/webr.qmd >}}
:::
:::::::::

{{< include ../../_utils/save_results.qmd >}}

{{< include ../../repro.qmd >}}